---
title: "10x-data-plot-dimreduction"
author: "Genevieve Housman"
date: "2/1/2020"
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## 10X Data Cellranger Analysis - Data Plotting

Plot aspects of 10X data.

Data being examined comes from 6 humans and 6 chimpanzees.
- GHO-1:  H1-I + C1-I (iPSCs from H23555 and C8861)
- GHO-2:  H1-M + C1-M (iPSC-derived MSCs from H23555 and C8861)
- GHO-4:  H1-O + C1-O (iPSC-MSC-derived osteoblasts from H23555 and C8861)
- HOU-10: H1-I-r2 + C1-I-r2 (iPSCs from H23555 and C8861)
- HOU-12: H1-M-r2 + C1-M-r2 (iPSC-derived MSCs from H23555 and C8861)
- HOU-16: H1-O-r2 + C1-O-r2 (iPSC-MSC-derived osteoblasts from H23555 and C8861)
- HOU-1:  H2-I + C2-I (iPSCs from H20157 and C3647)
- HOU-2:  H2-M + C2-M (iPSC-derived MSCs from H20157 and C3647)
- HOU-4:  H2-O + C2-O (iPSC-MSC-derived osteoblasts from H20157 and C3647)
- HOU-9:  H3-I + C3-I (iPSCs from H28126 and C3649)
- HOU-11: H3-M + C3-M (iPSC-derived MSCs from H28126 and C3649)
- HOU-15: H3-O + C3-O (iPSC-MSC-derived osteoblasts from H28126 and C3649)
- HOU-5:  H4-I + C4-I (iPSCs from H28834 and C40210)
- HOU-6:  H4-M + C4-M (iPSC-derived MSCs from H28834 and C40210)
- HOU-8:  H4-O + C4-O (iPSC-MSC-derived osteoblasts from H28834 and C40210)
- HOU-17:	H5-I + C5-I (iPSCs from H21792 and C40280)
- HOU-19:	H5-M + C5-M (iPSC-derived MSCs from H21792 and C40280)
- HOU-21:	H5-O + C5-O (iPSC-MSC-derived osteoblasts from H21792 and C40280)
- HOU-18:	H6-I + C6-I (iPSCs from H20961 and C3624)
- HOU-20:	H6-M + C6-M (iPSC-derived MSCs from H20961 and C3624)
- HOU-22:	H6-O + C6-O (iPSC-MSC-derived osteoblasts from H20961 and C3624)

Each dataset being evaluated comprises gene count data that were seperately calculated using both the hg38 human genome and the panTro6 chimpanzee genome, as well as the corresponding ortho-exon annotation in cellranger.Species-specific assignments were determined using human-chimpanzee genome combination with ortho-exon annotation.

```{r plots of different dimensional reductions}

#Load libraries
library(Seurat)
library(dplyr)
library(stringi)
library(stringr)
library(ggplot2)
library(colorspace)
library(RColorBrewer)

#Define main directory
dir <- '/project2/gilad/ghousman/skeletal-human-chimp/'

#Load batch info
batch <- read.csv(file=paste0(dir,'human-chimp-skeletal-scRNA/data/scrna-batch.csv'), header=TRUE, sep=",")

#Load data
#data <- readRDS(paste0(dir,"/human-chimp-skeletal-scRNA/data/cellranger-data-full/data.separate-indv.rds"))           #merge - total
#data <- readRDS(paste0(dir,"/human-chimp-skeletal-scRNA/data/cellranger-data-full/data.separate.rds"))                #merge - total
#data <- readRDS(paste0(dir,"/human-chimp-skeletal-scRNA/data/cellranger-data-full/data.merge.l.rds"))                 #merge - total
#data <- readRDS(paste0(dir,"/human-chimp-skeletal-scRNA/data/cellranger-data-full/data.merge.l.ipsc.rds"))            #merge - ipsc
#data <- readRDS(paste0(dir,"/human-chimp-skeletal-scRNA/data/cellranger-data-full/data.merge.l.msc.rds"))             #merge - msc
#data <- readRDS(paste0(dir,"/human-chimp-skeletal-scRNA/data/cellranger-data-full/data.merge.l.osteo.rds"))           #merge - osteo
#data <- readRDS(paste0(dir,"/human-chimp-skeletal-scRNA/data/cellranger-data-full/data.merge-6k.l.rds"))              #merge-6k - total
#data <- readRDS(paste0(dir,"/human-chimp-skeletal-scRNA/data/cellranger-data-full/data.merge-6k.l.ipsc.rds"))         #merge-6k - ipsc
#data <- readRDS(paste0(dir,"/human-chimp-skeletal-scRNA/data/cellranger-data-full/data.merge-6k.l.msc.rds"))          #merge-6k - msc
#data <- readRDS(paste0(dir,"/human-chimp-skeletal-scRNA/data/cellranger-data-full/data.merge-6k.l.osteo.rds"))        #merge-6k - osteo
#spps <- readRDS(paste0(dir,"/human-chimp-skeletal-scRNA/data/cellranger-data-full/data.spps.log.int.rds"))            #integrate across species - total
#spps <- readRDS(paste0(dir,"/human-chimp-skeletal-scRNA/data/cellranger-data-full/data.spps-cell.log.int.ipsc.rds"))  #integrate across species - ipsc
#spps <- readRDS(paste0(dir,"/human-chimp-skeletal-scRNA/data/cellranger-data-full/data.spps-cell.log.int.msc.rds"))   #integrate across species - msc
#spps <- readRDS(paste0(dir,"/human-chimp-skeletal-scRNA/data/cellranger-data-full/data.spps-cell.log.int.osteo.rds")) #integrate across species - osteo
#pair <- readRDS(paste0(dir,"/human-chimp-skeletal-scRNA/data/cellranger-data-full/data.pair.log.int.rds"))            #integrate across human-chimp pairs - total
#pair <- readRDS(paste0(dir,"/human-chimp-skeletal-scRNA/data/cellranger-data-full/data.pair-cell.log.int.ipsc.rds"))  #integrate across human-chimp pairs - ipsc
#pair <- readRDS(paste0(dir,"/human-chimp-skeletal-scRNA/data/cellranger-data-full/data.pair-cell.log.int.msc.rds"))   #integrate across human-chimp pairs - msc
#pair <- readRDS(paste0(dir,"/human-chimp-skeletal-scRNA/data/cellranger-data-full/data.pair-cell.log.int.osteo.rds")) #integrate across human-chimp pairs - osteo
indv <- readRDS(paste0(dir,"/human-chimp-skeletal-scRNA/data/cellranger-data-full/data.indv.log.int.rds"))            #integrate across individuals - total
#indv <- readRDS(paste0(dir,"/human-chimp-skeletal-scRNA/data/cellranger-data-full/data.indv-cell.log.int.ipsc.rds"))  #integrate across individuals - ipsc
#indv <- readRDS(paste0(dir,"/human-chimp-skeletal-scRNA/data/cellranger-data-full/data.indv-cell.log.int.msc.rds"))   #integrate across individuals - msc
#indv <- readRDS(paste0(dir,"/human-chimp-skeletal-scRNA/data/cellranger-data-full/data.indv-cell.log.int.osteo.rds")) #integrate across individuals - osteo

#Identify Clusters (increased resolution identifies greater number of clusters)
#for (i in 1:length(data)) {
#  print(i)
#  data[[i]] <- FindNeighbors(data[[i]], k.param=20, dims=1:100)
#  data[[i]] <- FindClusters(data[[i]], resolution=0.4)
#}
#data <- FindNeighbors(data, k.param=20, dims=1:100)
#data <- FindClusters(data, resolution=0.4)
#spps <- FindNeighbors(spps, k.param=20, dims=1:100)
#spps <- FindClusters(spps, resolution=0.4)
#pair <- FindNeighbors(pair, k.param=20, dims=1:100)
#pair <- FindClusters(pair, resolution=0.4)
indv <- FindNeighbors(indv, k.param=20, dims=1:100)
indv <- FindClusters(indv, resolution=0.4)

#data.separate-indv:: res=0.4: H1I/H1M/H1O/H1I2/H1M2/H1O2/H2I/H2M/H2O/H3I/H3M/H3O/H4I/H4M/H4O/H5I/H5M/H5O/H6I/H6M/H6O/
#                                 6/ 23/  5/   6/   6/   4/  7/ 33/  7/  4/  4/  2/  7/  4/ 10/  4/  5/  6/  7/ 10/  7/
#                              C1I/C1M/C1O/C1I2/C1M2/C1O2/C2I/C2M/C2O/C3I/C3M/C3O/C4I/C4M/C4O/C5I/C5M/C5O/C6I/C6M/C6O/
#                                5/  3/  3/   6/   8/   5/  6/  3/  4/  9/ 14/  6/  6/  3/  4/  7/ 33/  7/  6/  6/  5
#data.separate-pair:: res=0.4: H1C1I/H1C1M/H1C1O/H1C1I2/H1C1M2/H1C1O2/H2C2I/H2C2M/H2C2O/H3C3I/H3C3M/H3C3O/H4C4I/H4C4M/H4C4O/H5C5I/H5C5M/H5C5O/HC66I/H6C6M/H6C6O/
#                                 15/    9/    6/    15/     7/     6/    10/   7/   10/   10/    8/    5/   10/    8/    6/   11/    8/    5/   10/   10/   10
#merge1k::            res=0.4: ipsc - 22 clusters / msc - 22 clusters / osteo - 18 clusters / total - 38 clusters
#merge3k::            res=0.4: ipsc - 23 clusters / msc - 23 clusters / osteo - 24 clusters / total - 42 clusters
#merge6k::            res=0.4: ipsc - 24 clusters / msc - 22 clusters / osteo - 25 clusters / total - 46 clusters
#merge10k::           res=0.4: ipsc - 23 clusters / msc - 23 clusters / osteo - 27 clusters / total - 43 clusters
#merge-6k(spp)::      res=0.4: ipsc - 21 clusters / msc - 23 clusters / osteo - 25 clusters / total - 43 clusters
#spps1k::             res=0.4: ipsc - 15 clusters / msc - 20 clusters / osteo - 18 clusters / total - 33 clusters
#spps3k::             res=0.4: ipsc - 21 clusters / msc - 22 clusters / osteo - 19 clusters / total - 39 clusters
#spps6k::             res=0.4: ipsc - 22 clusters / msc - 24 clusters / osteo - 22 clusters / total - 39 clusters
#pair1k::             res=0.4: ipsc - 12 clusters / msc - 10 clusters / osteo - 11 clusters / total - 22 clusters
#pair3k::             res=0.4: ipsc - 20 clusters / msc - 17 clusters / osteo - 15 clusters / total - 26 clusters
#pair6k::             res=0.4: ipsc - 31 clusters / msc - 18 clusters / osteo - 19 clusters / total - 26 clusters
#indv1k::             res=0.4: ipsc - 13 clusters / msc -  8 clusters / osteo -  8 clusters / total - 15 clusters
#indv3k::             res=0.4: ipsc - 16 clusters / msc - 13 clusters / osteo - 12 clusters / total - 23 clusters
#indv6k::             res=0.4: ipsc - 24 clusters / msc - 17 clusters / osteo - 16 clusters / total - 20 clusters

#Plot Dimensional Reductions
for (i in 1:length(data)) {
  print(CombinePlots(plots=list((DimPlot(data[[i]], group.by="Cell.Type", reduction="umap") + labs(x="UMAP1",y="UMAP2")),
                                (DimPlot(data[[i]], group.by="Individual", reduction="umap") + labs(x="UMAP1",y="UMAP2")),
                                (DimPlot(data[[i]], group.by="seurat_clusters", reduction="umap") + labs(x="UMAP1",y="UMAP2"))), ncol=3))
}

DimPlot(data, group.by="Cell.Type", reduction="pca") +
  labs(x="PC1",y="PC2")
DimPlot(data, group.by="species", reduction="pca") +
  labs(x="PC1",y="PC2")
DimPlot(data, group.by="Individual", reduction="pca") +
  labs(x="PC1",y="PC2")
DimPlot(data, group.by="Cell.Type", reduction="umap") +
  labs(x="UMAP1",y="UMAP2")
DimPlot(data, group.by="species", reduction="umap") +
  labs(x="UMAP1",y="UMAP2")
DimPlot(data, group.by="Individual", reduction="umap") +
  labs(x="UMAP1",y="UMAP2")
DimPlot(data, group.by="seurat_clusters", reduction="umap") +
  labs(x="UMAP1",y="UMAP2")

DimPlot(spps, group.by="Cell.Type", reduction="pca") +
  labs(x="PC1",y="PC2")
DimPlot(spps, group.by="species", reduction="pca") +
  labs(x="PC1",y="PC2")
DimPlot(spps, group.by="Individual", reduction="pca") +
  labs(x="PC1",y="PC2")
DimPlot(spps, group.by="Cell.Type", reduction="umap") +
  labs(x="UMAP1",y="UMAP2")
DimPlot(spps, group.by="species", reduction="umap") +
  labs(x="UMAP1",y="UMAP2")
DimPlot(spps, group.by="Individual", reduction="umap") +
  labs(x="UMAP1",y="UMAP2")
DimPlot(spps, group.by="seurat_clusters", reduction="umap") +
  labs(x="UMAP1",y="UMAP2")

DimPlot(pair, group.by="Cell.Type", reduction="pca") +
  labs(x="PC1",y="PC2")
DimPlot(pair, group.by="species", reduction="pca") +
  labs(x="PC1",y="PC2")
DimPlot(pair, group.by="Individual", reduction="pca") +
  labs(x="PC1",y="PC2")
DimPlot(pair, group.by="Cell.Type", reduction="umap") +
  labs(x="UMAP1",y="UMAP2")
DimPlot(pair, group.by="species", reduction="umap") +
  labs(x="UMAP1",y="UMAP2")
DimPlot(pair, group.by="Individual", reduction="umap") +
  labs(x="UMAP1",y="UMAP2")
DimPlot(pair, group.by="seurat_clusters", reduction="umap") +
  labs(x="UMAP1",y="UMAP2")

DimPlot(indv, group.by="Cell.Type", reduction="pca") +
  labs(x="PC1",y="PC2")
DimPlot(indv, group.by="species", reduction="pca") +
  labs(x="PC1",y="PC2")
DimPlot(indv, group.by="Individual", reduction="pca") +
  labs(x="PC1",y="PC2")
DimPlot(indv, group.by="Cell.Type", reduction="umap") +
  labs(x="UMAP1",y="UMAP2")
DimPlot(indv, group.by="species", reduction="umap") +
  labs(x="UMAP1",y="UMAP2")
DimPlot(indv, group.by="Individual", reduction="umap") +
  labs(x="UMAP1",y="UMAP2")
DimPlot(indv, group.by="seurat_clusters", reduction="umap") +
  labs(x="UMAP1",y="UMAP2")

```

```{r plots of gene expression patterns - indv data only}

#Load libraries
library(Seurat)
library(dplyr)
library(stringi)
library(stringr)
library(ggplot2)
library(colorspace)
library(RColorBrewer)

#Define main directory
dir <- '/project2/gilad/ghousman/skeletal-human-chimp/'

#Load batch info
batch <- read.csv(file=paste0(dir,'human-chimp-skeletal-scRNA/data/scrna-batch.csv'), header=TRUE, sep=",")

#Load data
tot <- readRDS(paste0(dir,"/human-chimp-skeletal-scRNA/data/cellranger-data-full/data.indv.log.int.rds"))            #integrate across individuals - total
ips <- readRDS(paste0(dir,"/human-chimp-skeletal-scRNA/data/cellranger-data-full/data.indv-cell.log.int.ipsc.rds"))  #integrate across individuals - ipsc
msc <- readRDS(paste0(dir,"/human-chimp-skeletal-scRNA/data/cellranger-data-full/data.indv-cell.log.int.msc.rds"))   #integrate across individuals - msc
ost <- readRDS(paste0(dir,"/human-chimp-skeletal-scRNA/data/cellranger-data-full/data.indv-cell.log.int.osteo.rds")) #integrate across individuals - osteo

#Plot dimensional reductions
DimPlot(ips, group.by="Cell.Type", reduction="umap", pt.size=0.01, cols="pink") +
  labs(x="UMAP1",y="UMAP2")
DimPlot(msc, group.by="Cell.Type", reduction="umap", pt.size=0.01, cols="mediumorchid") +
  labs(x="UMAP1",y="UMAP2")
DimPlot(ost, group.by="Cell.Type", reduction="umap", pt.size=0.01, cols="dodgerblue") +
  labs(x="UMAP1",y="UMAP2")
DimPlot(tot, group.by="Cell.Type", reduction="umap", pt.size=0.01, cols=c("pink","mediumorchid","dodgerblue")) +
  labs(x="UMAP1",y="UMAP2")

DimPlot(ips, group.by="species", reduction="umap", pt.size=0.01, cols=c("grey70","grey30")) +
  labs(x="UMAP1",y="UMAP2")
DimPlot(msc, group.by="species", reduction="umap", pt.size=0.01, cols=c("grey70","grey30")) +
  labs(x="UMAP1",y="UMAP2")
DimPlot(ost, group.by="species", reduction="umap", pt.size=0.01, cols=c("grey70","grey30")) +
  labs(x="UMAP1",y="UMAP2")
DimPlot(tot, group.by="species", reduction="umap", pt.size=0.01, cols=c("grey70","grey30")) +
  labs(x="UMAP1",y="UMAP2")

DimPlot(ips, group.by="Individual", reduction="umap", pt.size=0.01) +
  labs(x="UMAP1",y="UMAP2")
DimPlot(msc, group.by="Individual", reduction="umap", pt.size=0.01) +
  labs(x="UMAP1",y="UMAP2")
DimPlot(ost, group.by="Individual", reduction="umap", pt.size=0.01) +
  labs(x="UMAP1",y="UMAP2")
DimPlot(tot, group.by="Individual", reduction="umap", pt.size=0.01) +
  labs(x="UMAP1",y="UMAP2")

#UMI and features counts
VlnPlot(tot, 'nCount_RNA', group.by='Cell.Type', split.by='species', pt.size=0, cols=c("grey70","grey30")) +
  theme(axis.text.x=element_text(angle=0, hjust=0.5)) +
  labs(x="",y="Number of UMI")
VlnPlot(tot, 'nFeature_RNA', group.by='Cell.Type', split.by='species', pt.size=0, cols=c("grey70","grey30")) +
  theme(axis.text.x=element_text(angle=0, hjust=0.5)) +
  labs(x="",y="Number of Genes")

CombinePlots(plots=list((VlnPlot(subset(tot,subset=species=="Human"),'nCount_RNA',group.by='Collection',
                                 pt.size=0, cols=rep(c("pink","mediumorchid","dodgerblue"),7)) +
                           theme(axis.text.x=element_text(angle=0, hjust=0.5), legend.position="none") +
                           labs(x="",y="Number of UMI") +
                           scale_x_discrete(labels=c("1"="H1\niPSCs","2"="H1\nMSCs","3"="H1\nOsteoblasts",
                                                     "4"="H1r2\niPSCs","5"="H1r2\nMSCs","6"="H1r2\nOsteoblasts",
                                                     "7"="H2\niPSCs","8"="H2\nMSCs","9"="H2\nOsteoblasts",
                                                     "10"="H3\niPSCs","11"="H3\nMSCs","12"="H3\nOsteoblasts",
                                                     "13"="H4\niPSCs","14"="H4\nMSCs","15"="H4\nOsteoblasts",
                                                     "16"="H5\niPSCs","17"="H5\nMSCs","18"="H5\nOsteoblasts",
                                                     "19"="H6\niPSCs","20"="H6\nMSCs","21"="H6\nOsteoblasts"))),
                        (VlnPlot(subset(tot,subset=species=="Chimp"),'nCount_RNA',group.by='Collection',
                                 pt.size=0, cols=rep(c("pink","mediumorchid","dodgerblue"),7)) +
                           theme(axis.text.x=element_text(angle=0, hjust=0.5), legend.position="none") +
                           labs(x="",y="Number of UMI") +
                           scale_x_discrete(labels=c("1"="C1\niPSCs","2"="C1\nMSCs","3"="C1\nOsteoblasts",
                                                     "4"="C1r2\niPSCs","5"="C1r2\nMSCs","6"="C1r2\nOsteoblasts",
                                                     "7"="C2\niPSCs","8"="C2\nMSCs","9"="C2\nOsteoblasts",
                                                     "10"="C3\niPSCs","11"="C3\nMSCs","12"="C3\nOsteoblasts",
                                                     "13"="C4\niPSCs","14"="C4\nMSCs","15"="C4\nOsteoblasts",
                                                     "16"="C5\niPSCs","17"="C5\nMSCs","18"="C5\nOsteoblasts",
                                                     "19"="C6\niPSCs","20"="C6\nMSCs","21"="C6\nOsteoblasts")))),
             ncol=1)

CombinePlots(plots=list((VlnPlot(subset(tot,subset=species=="Human"),'nFeature_RNA',group.by='Collection',
                                 pt.size=0, cols=rep(c("pink","mediumorchid","dodgerblue"),7)) +
                           theme(axis.text.x=element_text(angle=0, hjust=0.5), legend.position="none") +
                           labs(x="",y="Number of Genes") +
                           scale_x_discrete(labels=c("1"="H1\niPSCs","2"="H1\nMSCs","3"="H1\nOsteoblasts",
                                                     "4"="H1r2\niPSCs","5"="H1r2\nMSCs","6"="H1r2\nOsteoblasts",
                                                     "7"="H2\niPSCs","8"="H2\nMSCs","9"="H2\nOsteoblasts",
                                                     "10"="H3\niPSCs","11"="H3\nMSCs","12"="H3\nOsteoblasts",
                                                     "13"="H4\niPSCs","14"="H4\nMSCs","15"="H4\nOsteoblasts",
                                                     "16"="H5\niPSCs","17"="H5\nMSCs","18"="H5\nOsteoblasts",
                                                     "19"="H6\niPSCs","20"="H6\nMSCs","21"="H6\nOsteoblasts"))),
                        (VlnPlot(subset(tot,subset=species=="Chimp"),'nFeature_RNA',group.by='Collection',
                                 pt.size=0, cols=rep(c("pink","mediumorchid","dodgerblue"),7)) +
                           theme(axis.text.x=element_text(angle=0, hjust=0.5), legend.position="none") +
                           labs(x="",y="Number of Genes") +
                           scale_x_discrete(labels=c("1"="C1\niPSCs","2"="C1\nMSCs","3"="C1\nOsteoblasts",
                                                     "4"="C1r2\niPSCs","5"="C1r2\nMSCs","6"="C1r2\nOsteoblasts",
                                                     "7"="C2\niPSCs","8"="C2\nMSCs","9"="C2\nOsteoblasts",
                                                     "10"="C3\niPSCs","11"="C3\nMSCs","12"="C3\nOsteoblasts",
                                                     "13"="C4\niPSCs","14"="C4\nMSCs","15"="C4\nOsteoblasts",
                                                     "16"="C5\niPSCs","17"="C5\nMSCs","18"="C5\nOsteoblasts",
                                                     "19"="C6\niPSCs","20"="C6\nMSCs","21"="C6\nOsteoblasts")))),
             ncol=1)

VlnPlot(tot,'nFeature_RNA',group.by='Collection', split.by='species', pt.size=0, cols=c("grey70","grey30")) +
  theme(axis.text.x=element_text(angle=0, hjust=0.5)) +
  labs(x="",y="Number of Genes") +
  scale_x_discrete(labels=c("1"="H1+C1\niPSCs","2"="H1+C1\nMSCs","3"="H1+C1\nOsteoblasts",
                            "4"="H1r2+C1r2\niPSCs","5"="H1r2+C1r2\nMSCs","6"="H1r2+C1r2\nOsteoblasts",
                            "7"="H2+C2\niPSCs","8"="H2+C2\nMSCs","9"="H2+C2\nOsteoblasts",
                            "10"="H3+C3\niPSCs","11"="H3+C3\nMSCs","12"="H3+C3\nOsteoblasts",
                            "13"="H4+C4\niPSCs","14"="H4+C4\nMSCs","15"="H4+C4\nOsteoblasts",
                            "16"="H5+C5\niPSCs","17"="H5+C5\nMSCs","18"="H5+C5\nOsteoblasts",
                            "19"="H6+C6\niPSCs","20"="H6+C6\nMSCs","21"="H6+C6\nOsteoblasts"))
VlnPlot(tot,'nCount_RNA',group.by='Collection', split.by='species', pt.size=0, cols=c("grey70","grey30")) +
  theme(axis.text.x=element_text(angle=0, hjust=0.5)) +
  labs(x="",y="Number of UMI") +
  scale_x_discrete(labels=c("1"="H1+C1\niPSCs","2"="H1+C1\nMSCs","3"="H1+C1\nOsteoblasts",
                            "4"="H1r2+C1r2\niPSCs","5"="H1r2+C1r2\nMSCs","6"="H1r2+C1r2\nOsteoblasts",
                            "7"="H2+C2\niPSCs","8"="H2+C2\nMSCs","9"="H2+C2\nOsteoblasts",
                            "10"="H3+C3\niPSCs","11"="H3+C3\nMSCs","12"="H3+C3\nOsteoblasts",
                            "13"="H4+C4\niPSCs","14"="H4+C4\nMSCs","15"="H4+C4\nOsteoblasts",
                            "16"="H5+C5\niPSCs","17"="H5+C5\nMSCs","18"="H5+C5\nOsteoblasts",
                            "19"="H6+C6\niPSCs","20"="H6+C6\nMSCs","21"="H6+C6\nOsteoblasts"))

mean(tot@meta.data$nFeature_RNA[which(tot[['species']]=="Human")])        #3448.74
mean(tot@meta.data$nFeature_RNA[which(tot[['species']]=="Chimp")])        #3259.42
mean(tot@meta.data$nFeature_RNA[which(tot[['Cell.Type']]=="iPSC")])       #3454.82
mean(tot@meta.data$nFeature_RNA[which(tot[['Cell.Type']]=="MSC")])        #3769.14
mean(tot@meta.data$nFeature_RNA[which(tot[['Cell.Type']]=="Osteoblast")]) #2591.47

mean(tot@meta.data$nCount_RNA[which(tot[['species']]=="Human")])        #15444.44
mean(tot@meta.data$nCount_RNA[which(tot[['species']]=="Chimp")])        #12416.17
mean(tot@meta.data$nCount_RNA[which(tot[['Cell.Type']]=="iPSC")])       #13708.55
mean(tot@meta.data$nCount_RNA[which(tot[['Cell.Type']]=="MSC")])        #17142.95
mean(tot@meta.data$nCount_RNA[which(tot[['Cell.Type']]=="Osteoblast")]) # 9439.64

CombinePlots(plots=list((VlnPlot(subset(tot,subset=species=="Human"),'nFeature_RNA',group.by='Cell.Type',
                                 pt.size=0, cols=c("pink","mediumorchid","dodgerblue")) +
                           theme(axis.text.x=element_text(angle=0, hjust=0.5)) +
                           labs(x="",y="Number of Genes")),
                        (VlnPlot(subset(tot,subset=species=="Chimp"),'nFeature_RNA',group.by='Cell.Type',
                                 pt.size=0, cols=c("pink","mediumorchid","dodgerblue")) +
                           theme(axis.text.x=element_text(angle=0, hjust=0.5)) +
                           labs(x=""))),
                        ncol=2)

#pluripotency markers:  NANOG     (not in default search location)
#                       POU5F1
#                       TDGF1
#mesoderm markers:      MIXL1     (not in default search location)
#                       TBXT      (not in default search location)
#endomesoderm marker:   EOMES     (not in default search location)
#chondrocyte markers:   ACAN      (not in default search location)
#                       COL2A1
#                       SOX9      (not in default search location)
#                       COL10A1   (not in default search location)
#                       COL11A1
#                       SOX5      (not in default search location)
#                       SOX6      (not in default search location)
#osteoblast markers:    COL1A1
#                       BGLAP     (not in ortho-exon)
#                       RUNX2     (not in default search location)
#                       ALPL
#                       IBSP      (not in default search location)
#                       SP7       (not in default search location)
#adipocyte markers:     FABP4     (not in default search location)
#                       PPARG     (not in default search location)
#endoderm marker:       SOX17     (not in default search location)
#neuroectoderm marker:  SOX1      (not in default search location)
#neuroepithelial:       SOX2
#epithelial marker:     EPCAM
#hepatocyte marker:     ALB       (not in default search location)
#heart/NC marker:       HAND1     (not in default search location)
#cardiomyocyte marker:  TNNT2
#cariac muscle markers: MYL7
#                       MYH6      (not in default search location)
#neutral markers:       GAPDH     (not in default search location)
#                       GUSB      (not in default search location)
#                       YWHAZ     (not in default search location)

#Should be positive in iPSC/MSC: THY1 (CD90)
#Should be positive in MSC: ENG (CD105), NT5E (CD73)
#Can also be positive in MSC: CD44, ITGB1 (CD29), CD200, ALCAM (CD166), MCAM (CD146), NGFR (CD271)
#Should be negative in MSCs: CD34, PTPRC (CD45), ITGAM (CD11b), CD19, HLA-DRB1 (HLA-DR)
#Should also be negative in MSCs: CD14, CD79A (CD79α)

#Gene violin plots
c("NANOG","TDGF1","POU5F1","EOMES","THY1","MIXL1","TBXT","ENG","NT5E","CD44","ITGB1","CD200","ALCAM","MCAM","NGFR",
  "COL1A1","RUNX2","ALPL","IBSP","SP7","FABP4","PPARG","ACAN","COL2A1","SOX9","COL10A1","COL11A1","SOX5","SOX6",
  "SOX17","SOX1","SOX2","EPCAM","ALB","HAND1","TNNT2","MYL7","MYH6",
  "GAPDH","GUSB","YWHAZ","CD34","PTPRC","ITGAM","CD19","HLA-DRB1","CD14","CD79A")
FeaturePlot(tot, features="COL1A1", reduction="umap", cols=c("white","black"), pt.size=0.5, order=TRUE) + labs(x="UMAP1",y="UMAP2")
VlnPlot(tot, features="MYH6", group.by="Cell.Type", split.by="species", cols=c("grey70","grey30"), pt.size=0) +
  theme(axis.text.x=element_text(angle=0, hjust=0.5)) +
  labs(x="",y="Expression Level") +
  ylim(0,7)

#ipsc
FeaturePlot(tot, features="POU5F1", reduction="umap", cols=c("white","black"), pt.size=0.5, order=TRUE) + labs(x="UMAP1",y="UMAP2")
FeaturePlot(tot, features="TDGF1", reduction="umap", cols=c("white","black"), pt.size=0.5, order=TRUE) + labs(x="UMAP1",y="UMAP2")
VlnPlot(tot, features=c("POU5F1","TDGF1"), group.by="Cell.Type", cols=c("pink","mediumorchid","dodgerblue"), pt.size=0)
VlnPlot(tot, features="POU5F1", group.by="Cell.Type", cols=c("pink","mediumorchid","dodgerblue"), pt.size=0) +
  theme(axis.text.x=element_text(angle=0, hjust=0.5)) +
  labs(x="",y="Expression Level") +
  ylim(0,7)
VlnPlot(tot, features="POU5F1", group.by="Cell.Type", split.by="species", cols=c("grey70","grey30"), pt.size=0) +
  theme(axis.text.x=element_text(angle=0, hjust=0.5)) +
  labs(x="",y="Expression Level") +
  ylim(0,7)
VlnPlot(tot, features="TDGF1", group.by="Cell.Type", split.by="species", cols=c("grey70","grey30"), pt.size=0) +
  theme(axis.text.x=element_text(angle=0, hjust=0.5)) +
  labs(x="",y="Expression Level") +
  ylim(0,7)
#msc
FeaturePlot(tot, features="NT5E", reduction="umap", cols=c("grey90","black"), pt.size=0.5, order=TRUE) #CD73 - positive in MSC
FeaturePlot(tot, features="CD44", reduction="umap", cols=c("white","black"), pt.size=0.5, order=TRUE) + labs(x="UMAP1",y="UMAP2") #CD44 - can be positive
VlnPlot(tot, features=c("NT5E","CD44"), group.by="Cell.Type", cols=c("pink","mediumorchid","dodgerblue"), pt.size=0)
VlnPlot(tot, features="CD44", group.by="Cell.Type", cols=c("pink","mediumorchid","dodgerblue"), pt.size=0) +
  theme(axis.text.x=element_text(angle=0, hjust=0.5)) +
  labs(x="",y="Expression Level") +
  ylim(0,6)
VlnPlot(tot, features="CD44", group.by="Cell.Type", split.by="species", cols=c("grey70","grey30"), pt.size=0) +
  theme(axis.text.x=element_text(angle=0, hjust=0.5)) +
  labs(x="",y="Expression Level") +
  ylim(0,6)
#osteoblast
FeaturePlot(tot, features="COL1A1", cols=c("white","black"), pt.size=0.5, order=TRUE, reduction="umap") + labs(x="UMAP1",y="UMAP2")
FeaturePlot(tot, features="ALPL", cols=c("grey90","black"), pt.size=0.5, order=TRUE, reduction="umap")
VlnPlot(tot, features=c("COL1A1","ALPL"), group.by="Cell.Type", cols=c("pink","mediumorchid","dodgerblue"), pt.size=0)
VlnPlot(tot, features="COL1A1", group.by="Cell.Type", cols=c("pink","mediumorchid","dodgerblue"), pt.size=0) +
  theme(axis.text.x=element_text(angle=0, hjust=0.5)) +
  labs(x="",y="Expression Level") +
  ylim(0,8)
VlnPlot(tot, features="COL1A1", group.by="Cell.Type", split.by="species", cols=c("grey70","grey30"), pt.size=0) +
  theme(axis.text.x=element_text(angle=0, hjust=0.5)) +
  labs(x="",y="Expression Level") +
  ylim(0,8)

#Find differentially expressed genes
#find all markers distinguishing osteoblast clusters between humans and chimps
Idents(ost) <- factor(ost@meta.data$species, levels=c("Chimp","Human"))
o.tot <- FindMarkers(ost, ident.1="Human", ident.2="Chimp", logfc.threshold=0.25, test.use="wilcox", min.pct=0.25)
head(o.tot, n=5) #DKK2, CTHRC1, EMP2, RASL11A, PLAT
dim(o.tot) #25 genes total



#Identify Clusters (increased resolution identifies greater number of clusters)
tot <- FindNeighbors(tot, k.param=20, dims=1:100)
tot <- FindClusters(tot, resolution=0.4)          #15 clusters
ips <- FindNeighbors(ips, k.param=20, dims=1:100)
ips <- FindClusters(ips, resolution=0.4)          #13 clusters
msc <- FindNeighbors(msc, k.param=20, dims=1:100)
msc <- FindClusters(msc, resolution=0.4)          # 8 clusters
ost <- FindNeighbors(ost, k.param=20, dims=1:100)
ost <- FindClusters(ost, resolution=0.4)          # 8 clusters

DimPlot(tot, group.by="seurat_clusters", reduction="umap") +
  labs(x="UMAP1",y="UMAP2")

#Gene violin plots
c("NANOG","TDGF1","POU5F1","EOMES","THY1","MIXL1","TBXT","ENG","NT5E","CD44","ITGB1","CD200","ALCAM","MCAM","NGFR",
  "COL1A1","RUNX2","ALPL","IBSP","SP7","FABP4","PPARG","ACAN","COL2A1","SOX9","COL10A1","COL11A1","SOX5","SOX6","SOX17","SOX1","SOX2",
  "EPCAM","ALB","HAND1","TNNT2","MYL7","MYH6","GAPDH","GUSB","YWHAZ","CD34","PTPRC","ITGAM","CD19","HLA-DRB1","CD14","CD79A")
FeaturePlot(ost, features="COL1A1", reduction="umap", cols=c("white","black"), pt.size=0.5, order=TRUE) + labs(x="UMAP1",y="UMAP2")
VlnPlot(ost, features="COL1A1", group.by="seurat_clusters", split.by="species", cols=c("grey70","grey30"), pt.size=0) +
  theme(axis.text.x=element_text(angle=0, hjust=0.5)) +
  labs(x="",y="Expression Level")

#Find markers of clusters
ost.markers <- FindAllMarkers(ost, only.pos=TRUE, min.pct=0.25, logfc.threshold=0.25)
head(ost.markers, n=5)
top10 <- ost.markers %>% group_by(cluster) %>% top_n(n=10, wt=avg_logFC)
DoHeatmap(ost, features=top10$gene) + NoLegend()

FeaturePlot(ost, features=c("GDF15","TAGLN","PAPPA","TIMP1","ADAM33","COL3A1","RRM2"),
            reduction="umap", cols=c("white","black"), pt.size=0.5, order=TRUE)
VlnPlot(ost, features=c("GDF15","TAGLN","PAPPA","TIMP1","ADAM33","COL3A1","RRM2"),
        group.by="seurat_clusters", split.by="species", cols=c("grey70","grey30"), pt.size=0)

VlnPlot(ost, features=c("COL1A1","RUNX2","ALPL","IBSP","SP7"),
        group.by="seurat_clusters", split.by="species", cols=c("grey70","grey30"), pt.size=0)
VlnPlot(ost, features=c("FABP4","PPARG","ACAN","COL2A1","SOX9","COL10A1","COL11A1","SOX5","SOX6"),
        group.by="seurat_clusters", split.by="species", cols=c("grey70","grey30"), pt.size=0)


#Find differentially expressed genes - OSTEOBLASTS
#find all markers distinguishing osteoblast clusters between humans and chimps

ost0 <- subset(ost, seurat_clusters==0)
Idents(ost0) <- factor(ost0@meta.data$species, levels=c("Chimp","Human"))
o.0 <- FindMarkers(ost0, ident.1="Human", ident.2="Chimp", logfc.threshold=0.25, test.use="wilcox", min.pct=0.25)
head(o.0, n=5) #COL8A1, GDF15, THBS2, POSTN, IGFBP7
dim(o.0) #30 genes total

ost1 <- subset(ost, seurat_clusters==1)
Idents(ost1) <- factor(ost1@meta.data$species, levels=c("Chimp","Human"))
o.1 <- FindMarkers(ost1, ident.1="Human", ident.2="Chimp", logfc.threshold=0.25, test.use="wilcox", min.pct=0.25)
head(o.1, n=5) #EXPH5, PCDH10, EMP2, ITGA1, SERPINH1
dim(o.1) #77 genes total

ost2 <- subset(ost, seurat_clusters==2)
Idents(ost2) <- factor(ost2@meta.data$species, levels=c("Chimp","Human"))
o.2 <- FindMarkers(ost2, ident.1="Human", ident.2="Chimp", logfc.threshold=0.25, test.use="wilcox", min.pct=0.25)
head(o.2, n=5) #THBS2, COL8A1, SLC7A5, INHBA, POSTN
dim(o.2) #39 genes total

ost3 <- subset(ost, seurat_clusters==3)
Idents(ost3) <- factor(ost3@meta.data$species, levels=c("Chimp","Human"))
o.3 <- FindMarkers(ost3, ident.1="Human", ident.2="Chimp", logfc.threshold=0.25, test.use="wilcox", min.pct=0.25)
head(o.3, n=5) #DKK2, PMEPA1, SCG2, HIST2H2AC, COLEC12
dim(o.3) #81 genes total

ost4 <- subset(ost, seurat_clusters==4)
Idents(ost4) <- factor(ost4@meta.data$species, levels=c("Chimp","Human"))
o.4 <- FindMarkers(ost4, ident.1="Human", ident.2="Chimp", logfc.threshold=0.25, test.use="wilcox", min.pct=0.25)
head(o.4, n=5) #PLAT, DUSP23, HES4, HAPLN1, ENC1
dim(o.4) #137 genes total

ost5 <- subset(ost, seurat_clusters==5)
Idents(ost5) <- factor(ost5@meta.data$species, levels=c("Chimp","Human"))
o.5 <- FindMarkers(ost5, ident.1="Human", ident.2="Chimp", logfc.threshold=0.25, test.use="wilcox", min.pct=0.25)
head(o.5, n=5) #FGF2, FTH1, LGALS3, CDH11, CYSTM1
dim(o.5) #192 genes total

ost6 <- subset(ost, seurat_clusters==6)
Idents(ost6) <- factor(ost6@meta.data$species, levels=c("Chimp","Human"))
o.6 <- FindMarkers(ost6, ident.1="Human", ident.2="Chimp", logfc.threshold=0.25, test.use="wilcox", min.pct=0.25)
head(o.6, n=5) #CENPF, TOP2A, SC5D, TPM2, CXCL6
dim(o.6) #131 genes total

ost7 <- subset(ost, seurat_clusters==7)
Idents(ost7) <- factor(ost7@meta.data$species, levels=c("Chimp","Human"))
o.7 <- FindMarkers(ost7, ident.1="Human", ident.2="Chimp", logfc.threshold=0.25, test.use="wilcox", min.pct=0.25)
#no chimp cells in this cluster

#find all markers distinguishing humans and chimps across all cell types
o.tot <- intersect(rownames(o.0),
                   intersect(rownames(o.1),
                             intersect(rownames(o.2),
                                       intersect(rownames(o.3),
                                                 intersect(rownames(o.4),
                                                           intersect(rownames(o.5),rownames(o.6)))))))
library(VennDiagram)
overlap <- calculate.overlap( x <- list("cluster0"=rownames(o.0),
                                        "cluster1"=rownames(o.1),
                                        "cluster2"=rownames(o.2),
                                        "cluster3"=rownames(o.3),
                                        "cluster4"=rownames(o.4),
                                        "cluster5"=rownames(o.5),
                                        "cluster6"=rownames(o.6)))

rm(ost0,ost1,ost2,ost3,ost4,ost5,ost6,ost7)
rm(ost.markers,o.0,o.1,o.2,o.3,o.4,o.5,o.6,o.7,o.tot,top10)

#Find differentially expressed genes - IPSC
length(table(Idents(ips)))
x=0
while(x <= 12) {
  ipsX <- subset(ips, seurat_clusters==x)
  Idents(ipsX) <- factor(ipsX@meta.data$species, levels=c("Chimp","Human"))
  i.X <- FindMarkers(ipsX, ident.1="Human", ident.2="Chimp", logfc.threshold=0.25, test.use="wilcox", min.pct=0.25)
  print(x)
  print(head(i.X, n=5))
  print(dim(i.X))
  x=x+1
}
rm(x,ipsX,i.X)
#cluster  0: No features pass logfc.threshold threshold
#cluster  1: No features pass logfc.threshold threshold
#cluster  2: No features pass logfc.threshold threshold
#cluster  3:  13
#cluster  4: 222
#cluster  5:   8
#cluster  6: 105
#cluster  7:  25
#cluster  8:  66
#cluster  9: No features pass logfc.threshold threshold
#cluster 10: No features pass logfc.threshold threshold
#cluster 11: No features pass logfc.threshold threshold
#cluster 12: No features pass logfc.threshold threshold

#Find differentially expressed genes - MSC
length(table(Idents(msc)))
x=0
while(x <= 8) {
  mscX <- subset(msc, seurat_clusters==x)
  Idents(mscX) <- factor(mscX@meta.data$species, levels=c("Chimp","Human"))
  m.X <- FindMarkers(mscX, ident.1="Human", ident.2="Chimp", logfc.threshold=0.25, test.use="wilcox", min.pct=0.25)
  print(x)
  print(head(m.X, n=5))
  print(dim(m.X))
  x=x+1
}
rm(x,mscX,m.X)
#cluster  0:   4
#cluster  1:  22
#cluster  2:  20
#cluster  3:   4
#cluster  4:   4
#cluster  5:   2
#cluster  6:  92
#cluster  7: 265
#cluster  8: No features pass logfc.threshold threshold

#Find differentially expressed genes - TOTAL
length(table(Idents(tot)))
x=0
while(x <= 14) {
  totX <- subset(tot, seurat_clusters==x)
  Idents(totX) <- factor(totX@meta.data$species, levels=c("Chimp","Human"))
  t.X <- FindMarkers(totX, ident.1="Human", ident.2="Chimp", logfc.threshold=0.25, test.use="wilcox", min.pct=0.25)
  print(x)
  print(head(t.X, n=5))
  print(dim(t.X))
  x=x+1
}
rm(x,totX,t.X)
#cluster  0: 35
#cluster  1:  2
#cluster  2: 79
#cluster  3: 47
#cluster  4: No features pass logfc.threshold threshold
#cluster  5:  4
#cluster  6:  3
#cluster  7:  5
#cluster  8: 197
#cluster  9: 102
#cluster 10: 109
#cluster 11:  34
#cluster 12:  85
#cluster 13: 130
#cluster 14: No features pass logfc.threshold threshold

```
