---
title: "10x-data-qc"
author: "Genevieve Housman"
date: "Februrary 2021"
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# 10X Data Cellranger Analysis - Data QC

Examine QC metrics of processed 10X data using seurat.

```{r load libraries, message=FALSE}

library(Seurat)
library(dplyr)
library(stringi)
library(stringr)
library(ggplot2)
library(grid)
library(gridExtra)
library(colorspace)
library(RColorBrewer)

```

```{r load data}

#Load batch info
batch <- read.csv(file='./data/scrna-batch.csv', header=TRUE, sep=",")

#Read in files
data <- readRDS("./data/cellranger-data-full/data.rds")

```

### Data QC before barcode filtering

```{r, warning=FALSE}

pdf(file="./output/data-qc-prefilter.pdf", onefile=TRUE, width=21, height=21)

for (i in 1:length(data)) {
  
  collection <- paste0(data[[i]]@meta.data$Stage[1],": ",data[[i]]@meta.data$Pair[1])
  
  print(paste0("QC of ",collection))
  
  #Check and visualize numbers of cells per species
  print("Species Assignments")
  print(table(data[[i]]@meta.data$Species, useNA="always"))

  p1a <- ggplot(data[[i]]@meta.data, aes(x=Species)) +
           geom_bar() +
           labs(title=collection) +
           ylim(0,5000)
  
  #Check and visualize QC metrics
  # - number of umi per cell (count depth)
  # - number of genes per cell
  # - percent of mitochondrial reads (cell viability)
  # - expressed genes per cell (complexity)
  # - rarity of genes (cells expressing genes)
  
  #Not looking into expression of housekeeping genes at this time
  #https://broadinstitute.github.io/2019_scWorkshop/data-wrangling-scrnaseq.html
  
  #nCount_RNA: total number of molecules (umi) detected within a cell
  print("UMIs/Cell (Chimp and Human)")
  print(summary(data[[i]]@meta.data$nCount_RNA[which(data[[i]]@meta.data$Species=="Chimp")]))
  print(summary(data[[i]]@meta.data$nCount_RNA[which(data[[i]]@meta.data$Species=="Human")]))
  #nFeature_RNA: number of unique genes detected in each cell
  print("Genes/Cell (Chimp and Human)")
  print(summary(data[[i]]@meta.data$nFeature_RNA[which(data[[i]]@meta.data$Species=="Chimp")]))
  print(summary(data[[i]]@meta.data$nFeature_RNA[which(data[[i]]@meta.data$Species=="Human")]))
  #percent.mt: percentage of reads that map to the mitochondrial genome
  print("Percent.Mito/Cell (Chimp and Human)")
  print(summary(data[[i]]@meta.data$percent.mt[which(data[[i]]@meta.data$Species=="Chimp")]))
  print(summary(data[[i]]@meta.data$percent.mt[which(data[[i]]@meta.data$Species=="Human")]))
  #complexity
  data[[i]]@meta.data$Complexity <- log10(data[[i]]@meta.data$nCount_RNA/data[[i]]@meta.data$nFeature_RNA)
  print("Complexity (Chimp and Human)")
  print(summary(data[[i]]@meta.data$Complexity[which(data[[i]]@meta.data$Species=="Chimp")]))
  print(summary(data[[i]]@meta.data$Complexity[which(data[[i]]@meta.data$Species=="Human")]))
  
  p1b <- ggplot(data[[i]]@meta.data, aes(x=nCount_RNA,color=Species)) +
           geom_histogram(fill="white", alpha=0.1, position="identity", binwidth=100) +
           labs(title=collection,x="UMI per Cell",y="Count") +
           xlim(0,100000) +
           facet_grid(Species ~ .)
  p1c <- ggplot(data[[i]]@meta.data, aes(x=nFeature_RNA,color=Species)) +
           geom_histogram(fill="white", alpha=0.1, position="identity", binwidth=10) +
           labs(title=collection,x="Genes per Cell",y="Count") +
           xlim(0,10000) +
           facet_grid(Species ~ .)
  p1d <- ggplot(data[[i]]@meta.data, aes(x=percent.mt,color=Species)) +
           geom_histogram(fill="white", alpha=0.1, position="identity", binwidth=0.005) +
           labs(title=collection,x="Percent of Mitochondrial Reads",y="Count") +
           xlim(0,100) +
           facet_grid(Species ~ .)
    
  p2a <- ggplot(data[[i]]@meta.data, aes(x=Species, y=nCount_RNA, fill=Species)) +
           geom_violin() +
           geom_boxplot(fill="white", alpha=0.5, width=0.1) +
           labs(title=collection, x="Cell Type", y="UMI per Cell") +
           ylim(0,100000)
  p2b <- ggplot(data[[i]]@meta.data, aes(x=Species, y=nFeature_RNA, fill=Species)) +
           geom_violin() +
           geom_boxplot(fill="white", alpha=0.5, width=0.1) +
           labs(title=collection, x="Cell Type", y="Genes per Cell") +
           ylim(0,10000)
  p2c <- ggplot(data[[i]]@meta.data, aes(x=nCount_RNA, y=nFeature_RNA, color=Species)) +
           geom_point(size=2,alpha=0.3) +
           xlim(0,100000) +
           ylim(0,10000) +
           facet_grid(Species ~ .) +
           labs(title=collection,x="UMI per Cell",y="Genes per Cell")
  p2d <- ggplot(data[[i]]@meta.data, aes(x=Species, y=percent.mt, fill=Species)) +
           geom_violin() +
           geom_boxplot(fill="white", alpha=0.5, width=0.1, outlier.shape=NA) +
           labs(title=collection, x="Cell Type", y="Percent Mitochondrial Reads") +
           ylim(0,100)
  
  #possible threshold options: umi per cell > 1000 | genes per cell > 700 | percent mito < 25
  print("Species Assignments with Potential Filters Applied")
  print("[umi/cell > 1000 & < 60000 | gene/cell > 250 & < 7500 | percent.mt < 25]")
  print(table(data[[i]]@meta.data$Species[which(data[[i]]@meta.data$nCount_RNA > 1000 &
                                                data[[i]]@meta.data$nCount_RNA < 60000 &
                                                data[[i]]@meta.data$nFeature_RNA > 250 &
                                                data[[i]]@meta.data$nFeature_RNA < 7500 &
                                                data[[i]]@meta.data$percent.mt < 25)], useNA="always"))
  print("[percent.mt < 25]")
  print(table(data[[i]]@meta.data$Species[which(data[[i]]@meta.data$percent.mt < 25)], useNA="always"))

  p3a  <- ggplot(data[[i]]@meta.data, aes(x=Complexity,color=Species,fill=Species)) +
           geom_density(alpha=0.1, position="identity") +
           labs(title=collection,x="Complexity(log10GenesPerUMI)",y="Count") +
           xlim(0,2)
  p3b <- ggplot(data[[i]]@meta.data, aes(x=nCount_RNA,color=Species,fill=Species)) +
           geom_density(alpha=0.1, position="identity") +
           labs(title=collection,x="UMI per Cell",y="Count") +
           geom_vline(xintercept=c(1000,60000), linetype="dashed") +
           xlim(0,100000)
  p3c <- ggplot(data[[i]]@meta.data, aes(x=nFeature_RNA,color=Species,fill=Species)) +
           geom_density(alpha=0.1, position="identity") +
           labs(title=collection,x="Genes per Cell",y="Count") +
           geom_vline(xintercept=c(250,7500), linetype="dashed") +
           xlim(0,10000)
  p3d <- ggplot(data[[i]]@meta.data, aes(x=nCount_RNA, y=nFeature_RNA, color=percent.mt)) +
           geom_point(size=0.5, alpha=0.3) +
           scale_color_gradient2(low="black",mid="black",high="red",midpoint=25) +
           labs(title=collection,x="UMI per Cell",y="Genes per Cell") +
           xlim(0,100000) +
           ylim(0,10000) +
           geom_vline(xintercept=c(1000,60000), linetype="dashed") +
           geom_hline(yintercept=c(250,7500), linetype="dashed") +
           facet_grid(Species ~ .)

  #investigate potential that cells are actually failed libraries (lower end outliers) or are cell doublets (higher end outliers)
  p4a <- ggplot(subset(data[[i]]@meta.data,Species%in%"Chimp"),
                aes(x=1:length(nCount_RNA),y=sort(nCount_RNA))) +
           geom_point(size=0.5,alpha=0.5,color="#F8766D") +
           labs(title=collection, x="Cell Rank", y="UMI per Cell") +
           ylim(0,100000) +
           geom_hline(yintercept=c(1000,60000), linetype="dashed")
  p4b <- ggplot(subset(data[[i]]@meta.data,Species%in%"Human"),
                aes(x=1:length(nCount_RNA),y=sort(nCount_RNA))) +
           geom_point(size=0.5,alpha=0.5,color="#00BFC4") +
           labs(title=collection, x="Cell Rank", y="UMI per Cell") +
           ylim(0,100000) +
           geom_hline(yintercept=c(1000,60000), linetype="dashed")
  p4c <- ggplot(subset(data[[i]]@meta.data,Species%in%"Chimp"),
                aes(x=1:length(nFeature_RNA),y=sort(nFeature_RNA))) +
           geom_point(size=0.5,alpha=0.5,color="#F8766D") +
           labs(title=collection, x="Cell Rank", y="Genes per Cell") +
           ylim(0,10000) +
           geom_hline(yintercept=c(250,7500), linetype="dashed")
  p4d <- ggplot(subset(data[[i]]@meta.data,Species%in%"Human"),
                aes(x=1:length(nFeature_RNA),y=sort(nFeature_RNA))) +
           geom_point(size=0.5,alpha=0.5,color="#00BFC4") +
           labs(title=collection, x="Cell Rank", y="Genes per Cell") +
           ylim(0,10000) +
           geom_hline(yintercept=c(250,7500), linetype="dashed")
  
  #look at genes instead of cells
  #umis per gene & rarity of genes (cells expressing genes)
  #thresholds: umi per gene > 1
  df <- data.frame(
    counts_per_gene_chimp <- Matrix::rowSums(data[[i]]@assays$RNA@counts[,which(data[[i]]@meta.data$Species=="Chimp")]),
    counts_per_gene_human <- Matrix::rowSums(data[[i]]@assays$RNA@counts[,which(data[[i]]@meta.data$Species=="Human")]),
    cells_per_gene_chimp <- Matrix::rowSums(data[[i]]@assays$RNA@counts[,which(data[[i]]@meta.data$Species=="Chimp")]>0),
    cells_per_gene_human <- Matrix::rowSums(data[[i]]@assays$RNA@counts[,which(data[[i]]@meta.data$Species=="Human")]>0)
  )
  rm(counts_per_gene_chimp,counts_per_gene_human,cells_per_gene_chimp,cells_per_gene_human)
  colnames(df) <- c("counts_per_gene_chimp","counts_per_gene_human","cells_per_gene_chimp","cells_per_gene_human")

  print(paste0("Total Genes: ",length(rownames(data[[i]]@assays$RNA@counts))))
  print("Genes with at least one UMI (Chimp, Human, and Total)")
  print(table(Matrix::rowSums(data[[i]]@assays$RNA@counts[,which(data[[i]]@meta.data$Species=="Chimp")])>0))
  print(table(Matrix::rowSums(data[[i]]@assays$RNA@counts[,which(data[[i]]@meta.data$Species=="Human")])>0))
  print(table(Matrix::rowSums(data[[i]]@assays$RNA@counts)>0))

  p5a <- ggplot(df, aes(x=log10(counts_per_gene_chimp+1))) +
           geom_histogram(color="#F8766D", fill="white", alpha=0.1, binwidth=0.05) +
           ylim(0,6000) +
           geom_vline(xintercept=log10(1.5), linetype="dashed") +
           labs(title=collection, x="Log10 UMIs per Gene (Chimp)", y="Count")
  p5b <- ggplot(df, aes(x=log10(counts_per_gene_human+1))) +
           geom_histogram(color="#00BFC4", fill="white", alpha=0.1, binwidth=0.05) +
           ylim(0,6000) +
           geom_vline(xintercept=log10(1.5), linetype="dashed") +
           labs(title=collection, x="Log10 UMIs per Gene (Human)", y="Count")
  p5c <- ggplot(df, aes(x=log10(cells_per_gene_chimp+1))) +
           geom_histogram(color="#F8766D", fill="white", alpha=0.1, binwidth=0.05) +
           ylim(0,6000) +
           geom_vline(xintercept=log10(1.5), linetype="dashed") +
           labs(title=collection,x="Log10 Cells per Gene (Chimp)",y="Count")
  p5d <- ggplot(df, aes(x=log10(cells_per_gene_human+1))) +
           geom_histogram(color="#00BFC4", fill="white", alpha=0.1, binwidth=0.05) +
           ylim(0,6000) +
           geom_vline(xintercept=log10(1.5), linetype="dashed") +
           labs(title=collection,x="Log10 Cells per Gene (Human)",y="Count")
  
  grid.arrange(grobs=list(p1a,p1b,p1c,p1d,
                          p2a,p2b,p2c,p2d,
                          p3a,p3b,p3c,p3d,
                          p4a,p4b,p4c,p4d,
                          p5a,p5b,p5c,p5d), ncol=4)

  rm(p1a,p1b,p1c,p1d, p2a,p2b,p2c,p2d, p3a,p3b,p3c,p3d, p4a,p4b,p4c,p4d, p5a,p5b,p5c,p5d)
  
}

dev.off()

```

### Barcode filtering

External Resources:

* https://scrnaseq-course.cog.sanger.ac.uk/website/cleaning-the-expression-matrix.html
* https://broadinstitute.github.io/2019_scWorkshop/data-wrangling-scrnaseq.html#preprocessing-step-1-filter-out-low-quality-cells
* https://hbctraining.github.io/scRNA-seq/lessons/04_SC_quality_control.html

Filtering can be based on:

1. number of read counts in each cell (count depth)
2. number of unique genes detected in each cell (complexity)
3. detectable genes
     + detectable gene: >=1 UMI in >=2 cells (>=5 reads in >=2 cells)
     + genes must be filtered after cell filtering since some genes may only be detected in poor quality cells
     + notes: no filtering on genes at this time (should do this before constructing a seurat object)
4. mitochondrial RNA content
     + rationale: cells lyse more easily than mitochondria, so high mito RNA content can signal cell lysis and RNA degradation (ideally percent.mt should be <5%)

Based on the QC of my scRNA-seq data:

* some variability in reasonable UMIs/cell and genes/cell cutoff thresholds across collections
* percent.mt < 25 seems like a better threshold than percent.mt < 20/15/10 for retaining a good number of cells
* setting a cutoff for percent.mt < 25 filters out some cells but fewer than when adding UMI and gene cutoffs
* setting a cutoff for UMI count > 1000 & < 60000, gene count > 250 & < 7500, and percent.mt < 25 seems relatively good across datasets

Decided to filter cells using percent percent mitochondrial reads < 25 cutoff threshold as a lenient threshold and UMI counts per cell > 1000 & < 60000, gene counts per cell > 250 & < 750, and percent mitochondrial reads < 25 as a conservative threshold.

```{r, eval=FALSE}

#lenient threshold: percent.mt < 25

data.filter <- list()
i=1
while (i <= length(data)) {
  data.filter[[i]] <- subset(data[[i]], subset=percent.mt<25)
  i=i+1
}

#Save data
saveRDS(data.filter, file="./data/cellranger-data-full/data.filterL.rds")

```

```{r, eval=FALSE}

#conservative threshold: umi per cell > 1000 | genes per cell > 700 | percent mito < 25

data.filter <- list()
i=1
while (i <= length(data)) {
  data.filter[[i]] <- subset(data[[i]], subset=(percent.mt<25 & nCount_RNA>1000 & nCount_RNA<60000 & nFeature_RNA>250 & nFeature_RNA<7500))
  i=i+1
}

#Save data
saveRDS(data.filter, file="./data/cellranger-data-full/data.filterC.rds")

```

### Data QC after barcode filtering (lenient)

```{r, warning=FALSE, eval=FALSE}

data.filter <- readRDS("./data/cellranger-data-full/data.filterL.rds")

pdf(file="./output/data-qc-postfilterL.pdf", onefile=TRUE, width=21, height=21)

for (i in 1:length(data.filter)) {
  
  collection <- paste0(data.filter[[i]]@meta.data$Stage[1],": ",data.filter[[i]]@meta.data$Pair[1])
  
  print(paste0("QC of ",collection))
  
  #Check and visualize numbers of cells per species
  print("Species Assignments")
  print(table(data.filter[[i]]@meta.data$Species, useNA="always"))

  p1a <- ggplot(data.filter[[i]]@meta.data, aes(x=Species)) +
           geom_bar() +
           labs(title=collection) +
           ylim(0,5000)
  
  #Check and visualize QC metrics
  # - number of umi per cell (count depth)
  # - number of genes per cell
  # - percent of mitochondrial reads (cell viability)
  # - expressed genes per cell (complexity)
  # - rarity of genes (cells expressing genes)
  
  #Not looking into expression of housekeeping genes at this time
  #https://broadinstitute.github.io/2019_scWorkshop/data-wrangling-scrnaseq.html
  
  #nCount_RNA: total number of molecules (umi) detected within a cell
  print("UMIs/Cell (Chimp and Human)")
  print(summary(data.filter[[i]]@meta.data$nCount_RNA[which(data.filter[[i]]@meta.data$Species=="Chimp")]))
  print(summary(data.filter[[i]]@meta.data$nCount_RNA[which(data.filter[[i]]@meta.data$Species=="Human")]))
  #nFeature_RNA: number of unique genes detected in each cell
  print("Genes/Cell (Chimp and Human)")
  print(summary(data.filter[[i]]@meta.data$nFeature_RNA[which(data.filter[[i]]@meta.data$Species=="Chimp")]))
  print(summary(data.filter[[i]]@meta.data$nFeature_RNA[which(data.filter[[i]]@meta.data$Species=="Human")]))
  #percent.mt: percentage of reads that map to the mitochondrial genome
  print("Percent.Mito/Cell (Chimp and Human)")
  print(summary(data.filter[[i]]@meta.data$percent.mt[which(data.filter[[i]]@meta.data$Species=="Chimp")]))
  print(summary(data.filter[[i]]@meta.data$percent.mt[which(data.filter[[i]]@meta.data$Species=="Human")]))
  #complexity
  data[[i]]@meta.data$Complexity <- log10(data[[i]]@meta.data$nCount_RNA/data[[i]]@meta.data$nFeature_RNA)
  print("Complexity (Chimp and Human)")
  print(summary(data[[i]]@meta.data$Complexity[which(data[[i]]@meta.data$Species=="Chimp")]))
  print(summary(data[[i]]@meta.data$Complexity[which(data[[i]]@meta.data$Species=="Human")]))
  
  p1b <- ggplot(data[[i]]@meta.data, aes(x=nCount_RNA,color=Species)) +
           geom_histogram(fill="white", alpha=0.1, position="identity", binwidth=100) +
           labs(title=collection,x="UMI per Cell",y="Count") +
           xlim(0,100000) +
           facet_grid(Species ~ .)
  p1c <- ggplot(data[[i]]@meta.data, aes(x=nFeature_RNA,color=Species)) +
           geom_histogram(fill="white", alpha=0.1, position="identity", binwidth=10) +
           labs(title=collection,x="Genes per Cell",y="Count") +
           xlim(0,10000) +
           facet_grid(Species ~ .)
  p1d <- ggplot(data[[i]]@meta.data, aes(x=percent.mt,color=Species)) +
           geom_histogram(fill="white", alpha=0.1, position="identity", binwidth=0.005) +
           labs(title=collection,x="Percent of Mitochondrial Reads",y="Count") +
           xlim(0,100) +
           facet_grid(Species ~ .)
    
  p2a <- ggplot(data[[i]]@meta.data, aes(x=Species, y=nCount_RNA, fill=Species)) +
           geom_violin() +
           geom_boxplot(fill="white", alpha=0.5, width=0.1) +
           labs(title=collection, x="Cell Type", y="UMI per Cell") +
           ylim(0,100000)
  p2b <- ggplot(data[[i]]@meta.data, aes(x=Species, y=nFeature_RNA, fill=Species)) +
           geom_violin() +
           geom_boxplot(fill="white", alpha=0.5, width=0.1) +
           labs(title=collection, x="Cell Type", y="Genes per Cell") +
           ylim(0,10000)
  p2c <- ggplot(data[[i]]@meta.data, aes(x=nCount_RNA, y=nFeature_RNA, color=Species)) +
           geom_point(size=2,alpha=0.3) +
           xlim(0,100000) +
           ylim(0,10000) +
           facet_grid(Species ~ .) +
           labs(title=collection,x="UMI per Cell",y="Genes per Cell")
  p2d <- ggplot(data[[i]]@meta.data, aes(x=Species, y=percent.mt, fill=Species)) +
           geom_violin() +
           geom_boxplot(fill="white", alpha=0.5, width=0.1, outlier.shape=NA) +
           labs(title=collection, x="Cell Type", y="Percent Mitochondrial Reads") +
           ylim(0,100)

  p3a  <- ggplot(data[[i]]@meta.data, aes(x=Complexity,color=Species,fill=Species)) +
           geom_density(alpha=0.1, position="identity") +
           labs(title=collection,x="Complexity(log10GenesPerUMI)",y="Count") +
           xlim(0,2)
  p3b <- ggplot(data[[i]]@meta.data, aes(x=nCount_RNA,color=Species,fill=Species)) +
           geom_density(alpha=0.1, position="identity") +
           labs(title=collection,x="UMI per Cell",y="Count") +
           xlim(0,100000)
  p3c <- ggplot(data[[i]]@meta.data, aes(x=nFeature_RNA,color=Species,fill=Species)) +
           geom_density(alpha=0.1, position="identity") +
           labs(title=collection,x="Genes per Cell",y="Count") +
           xlim(0,10000)
  p3d <- ggplot(data[[i]]@meta.data, aes(x=nCount_RNA, y=nFeature_RNA, color=percent.mt)) +
           geom_point(size=0.5, alpha=0.3) +
           scale_color_gradient2(low="black",mid="black",high="red",midpoint=25) +
           labs(title=collection,x="UMI per Cell",y="Genes per Cell") +
           xlim(0,100000) +
           ylim(0,10000) +
           facet_grid(Species ~ .)

  #investigate potential that cells are actually failed libraries (lower end outliers) or are cell doublets (higher end outliers)
  p4a <- ggplot(subset(data[[i]]@meta.data,Species%in%"Chimp"),
                aes(x=1:length(nCount_RNA),y=sort(nCount_RNA))) +
           geom_point(size=0.5,alpha=0.5,color="#F8766D") +
           labs(title=collection, x="Cell Rank", y="UMI per Cell") +
           ylim(0,100000)
  p4b <- ggplot(subset(data[[i]]@meta.data,Species%in%"Human"),
                aes(x=1:length(nCount_RNA),y=sort(nCount_RNA))) +
           geom_point(size=0.5,alpha=0.5,color="#00BFC4") +
           labs(title=collection, x="Cell Rank", y="UMI per Cell") +
           ylim(0,100000)
  p4c <- ggplot(subset(data[[i]]@meta.data,Species%in%"Chimp"),
                aes(x=1:length(nFeature_RNA),y=sort(nFeature_RNA))) +
           geom_point(size=0.5,alpha=0.5,color="#F8766D") +
           labs(title=collection, x="Cell Rank", y="Genes per Cell") +
           ylim(0,10000)
  p4d <- ggplot(subset(data[[i]]@meta.data,Species%in%"Human"),
                aes(x=1:length(nFeature_RNA),y=sort(nFeature_RNA))) +
           geom_point(size=0.5,alpha=0.5,color="#00BFC4") +
           labs(title=collection, x="Cell Rank", y="Genes per Cell") +
           ylim(0,10000)
  
  #look at genes instead of cells
  #umis per gene & rarity of genes (cells expressing genes)
  #thresholds: umi per gene > 1
  df <- data.frame(
    counts_per_gene_chimp <- Matrix::rowSums(data.filter[[i]]@assays$RNA@counts[,which(data.filter[[i]]@meta.data$Species=="Chimp")]),
    counts_per_gene_human <- Matrix::rowSums(data.filter[[i]]@assays$RNA@counts[,which(data.filter[[i]]@meta.data$Species=="Human")]),
    cells_per_gene_chimp <- Matrix::rowSums(data.filter[[i]]@assays$RNA@counts[,which(data.filter[[i]]@meta.data$Species=="Chimp")]>0),
    cells_per_gene_human <- Matrix::rowSums(data.filter[[i]]@assays$RNA@counts[,which(data.filter[[i]]@meta.data$Species=="Human")]>0)
  )
  rm(counts_per_gene_chimp,counts_per_gene_human,cells_per_gene_chimp,cells_per_gene_human)
  colnames(df) <- c("counts_per_gene_chimp","counts_per_gene_human","cells_per_gene_chimp","cells_per_gene_human")

  print(paste0("Total Genes: ",length(rownames(data.filter[[i]]@assays$RNA@counts))))
  print("Genes with at least one UMI (Chimp, Human, and Total)")
  print(table(Matrix::rowSums(data.filter[[i]]@assays$RNA@counts[,which(data.filter[[i]]@meta.data$Species=="Chimp")])>0))
  print(table(Matrix::rowSums(data.filter[[i]]@assays$RNA@counts[,which(data.filter[[i]]@meta.data$Species=="Human")])>0))
  print(table(Matrix::rowSums(data.filter[[i]]@assays$RNA@counts)>0))
  
  p5a <- ggplot(df, aes(x=log10(counts_per_gene_chimp+1))) +
           geom_histogram(color="#F8766D", fill="white", alpha=0.1, binwidth=0.05) +
           ylim(0,6000) +
           geom_vline(xintercept=log10(1.5), linetype="dashed") +
           labs(title=collection, x="Log10 UMIs per Gene (Chimp)", y="Count")
  p5b <- ggplot(df, aes(x=log10(counts_per_gene_human+1))) +
           geom_histogram(color="#00BFC4", fill="white", alpha=0.1, binwidth=0.05) +
           ylim(0,6000) +
           geom_vline(xintercept=log10(1.5), linetype="dashed") +
           labs(title=collection, x="Log10 UMIs per Gene (Human)", y="Count")
  p5c <- ggplot(df, aes(x=log10(cells_per_gene_chimp+1))) +
           geom_histogram(color="#F8766D", fill="white", alpha=0.1, binwidth=0.05) +
           ylim(0,6000) +
           geom_vline(xintercept=log10(1.5), linetype="dashed") +
           labs(title=collection,x="Log10 Cells per Gene (Chimp)",y="Count")
  p5d <- ggplot(df, aes(x=log10(cells_per_gene_human+1))) +
           geom_histogram(color="#00BFC4", fill="white", alpha=0.1, binwidth=0.05) +
           ylim(0,6000) +
           geom_vline(xintercept=log10(1.5), linetype="dashed") +
           labs(title=collection,x="Log10 Cells per Gene (Human)",y="Count")

  grid.arrange(grobs=list(p1a,p1b,p1c,p1d,
                          p2a,p2b,p2c,p2d,
                          p3a,p3b,p3c,p3d,
                          p4a,p4b,p4c,p4d,
                          p5a,p5b,p5c,p5d), ncol=4)
  
  rm(p1a,p1b,p1c,p1d, p2a,p2b,p2c,p2d, p3a,p3b,p3c,p3d, p4a,p4b,p4c,p4d, p5a,p5b,p5c,p5d)
  
}

dev.off()

```

### Data QC after barcode filtering (conservative)

```{r, warning=FALSE, eval=FALSE}

data.filter <- readRDS("./data/cellranger-data-full/data.filterC.rds")

pdf(file="./output/data-qc-postfilterC.pdf", onefile=TRUE, width=21, height=21)

for (i in 1:length(data.filter)) {
  
  collection <- paste0(data.filter[[i]]@meta.data$Stage[1],": ",data.filter[[i]]@meta.data$Pair[1])
  
  print(paste0("QC of ",collection))
  
  #Check and visualize numbers of cells per species
  print("Species Assignments")
  print(table(data.filter[[i]]@meta.data$Species, useNA="always"))

  p1a <- ggplot(data.filter[[i]]@meta.data, aes(x=Species)) +
           geom_bar() +
           labs(title=collection) +
           ylim(0,5000)
  
  #Check and visualize QC metrics
  # - number of umi per cell (count depth)
  # - number of genes per cell
  # - percent of mitochondrial reads (cell viability)
  # - expressed genes per cell (complexity)
  # - rarity of genes (cells expressing genes)
  
  #Not looking into expression of housekeeping genes at this time
  #https://broadinstitute.github.io/2019_scWorkshop/data-wrangling-scrnaseq.html
  
  #nCount_RNA: total number of molecules (umi) detected within a cell
  print("UMIs/Cell (Chimp and Human)")
  print(summary(data.filter[[i]]@meta.data$nCount_RNA[which(data.filter[[i]]@meta.data$Species=="Chimp")]))
  print(summary(data.filter[[i]]@meta.data$nCount_RNA[which(data.filter[[i]]@meta.data$Species=="Human")]))
  #nFeature_RNA: number of unique genes detected in each cell
  print("Genes/Cell (Chimp and Human)")
  print(summary(data.filter[[i]]@meta.data$nFeature_RNA[which(data.filter[[i]]@meta.data$Species=="Chimp")]))
  print(summary(data.filter[[i]]@meta.data$nFeature_RNA[which(data.filter[[i]]@meta.data$Species=="Human")]))
  #percent.mt: percentage of reads that map to the mitochondrial genome
  print("Percent.Mito/Cell (Chimp and Human)")
  print(summary(data.filter[[i]]@meta.data$percent.mt[which(data.filter[[i]]@meta.data$Species=="Chimp")]))
  print(summary(data.filter[[i]]@meta.data$percent.mt[which(data.filter[[i]]@meta.data$Species=="Human")]))
  #complexity
  data[[i]]@meta.data$Complexity <- log10(data[[i]]@meta.data$nCount_RNA/data[[i]]@meta.data$nFeature_RNA)
  print("Complexity (Chimp and Human)")
  print(summary(data[[i]]@meta.data$Complexity[which(data[[i]]@meta.data$Species=="Chimp")]))
  print(summary(data[[i]]@meta.data$Complexity[which(data[[i]]@meta.data$Species=="Human")]))
  
  p1b <- ggplot(data[[i]]@meta.data, aes(x=nCount_RNA,color=Species)) +
           geom_histogram(fill="white", alpha=0.1, position="identity", binwidth=100) +
           labs(title=collection,x="UMI per Cell",y="Count") +
           xlim(0,100000) +
           facet_grid(Species ~ .)
  p1c <- ggplot(data[[i]]@meta.data, aes(x=nFeature_RNA,color=Species)) +
           geom_histogram(fill="white", alpha=0.1, position="identity", binwidth=10) +
           labs(title=collection,x="Genes per Cell",y="Count") +
           xlim(0,10000) +
           facet_grid(Species ~ .)
  p1d <- ggplot(data[[i]]@meta.data, aes(x=percent.mt,color=Species)) +
           geom_histogram(fill="white", alpha=0.1, position="identity", binwidth=0.005) +
           labs(title=collection,x="Percent of Mitochondrial Reads",y="Count") +
           xlim(0,100) +
           facet_grid(Species ~ .)
    
  p2a <- ggplot(data[[i]]@meta.data, aes(x=Species, y=nCount_RNA, fill=Species)) +
           geom_violin() +
           geom_boxplot(fill="white", alpha=0.5, width=0.1) +
           labs(title=collection, x="Cell Type", y="UMI per Cell") +
           ylim(0,100000)
  p2b <- ggplot(data[[i]]@meta.data, aes(x=Species, y=nFeature_RNA, fill=Species)) +
           geom_violin() +
           geom_boxplot(fill="white", alpha=0.5, width=0.1) +
           labs(title=collection, x="Cell Type", y="Genes per Cell") +
           ylim(0,10000)
  p2c <- ggplot(data[[i]]@meta.data, aes(x=nCount_RNA, y=nFeature_RNA, color=Species)) +
           geom_point(size=2,alpha=0.3) +
           xlim(0,100000) +
           ylim(0,10000) +
           facet_grid(Species ~ .) +
           labs(title=collection,x="UMI per Cell",y="Genes per Cell")
  p2d <- ggplot(data[[i]]@meta.data, aes(x=Species, y=percent.mt, fill=Species)) +
           geom_violin() +
           geom_boxplot(fill="white", alpha=0.5, width=0.1, outlier.shape=NA) +
           labs(title=collection, x="Cell Type", y="Percent Mitochondrial Reads") +
           ylim(0,100)

  p3a  <- ggplot(data[[i]]@meta.data, aes(x=Complexity,color=Species,fill=Species)) +
           geom_density(alpha=0.1, position="identity") +
           labs(title=collection,x="Complexity(log10GenesPerUMI)",y="Count") +
           xlim(0,2)
  p3b <- ggplot(data[[i]]@meta.data, aes(x=nCount_RNA,color=Species,fill=Species)) +
           geom_density(alpha=0.1, position="identity") +
           labs(title=collection,x="UMI per Cell",y="Count") +
           xlim(0,100000)
  p3c <- ggplot(data[[i]]@meta.data, aes(x=nFeature_RNA,color=Species,fill=Species)) +
           geom_density(alpha=0.1, position="identity") +
           labs(title=collection,x="Genes per Cell",y="Count") +
           xlim(0,10000)
  p3d <- ggplot(data[[i]]@meta.data, aes(x=nCount_RNA, y=nFeature_RNA, color=percent.mt)) +
           geom_point(size=0.5, alpha=0.3) +
           scale_color_gradient2(low="black",mid="black",high="red",midpoint=25) +
           labs(title=collection,x="UMI per Cell",y="Genes per Cell") +
           xlim(0,100000) +
           ylim(0,10000) +
           facet_grid(Species ~ .)

  #investigate potential that cells are actually failed libraries (lower end outliers) or are cell doublets (higher end outliers)
  p4a <- ggplot(subset(data[[i]]@meta.data,Species%in%"Chimp"),
                aes(x=1:length(nCount_RNA),y=sort(nCount_RNA))) +
           geom_point(size=0.5,alpha=0.5,color="#F8766D") +
           labs(title=collection, x="Cell Rank", y="UMI per Cell") +
           ylim(0,100000)
  p4b <- ggplot(subset(data[[i]]@meta.data,Species%in%"Human"),
                aes(x=1:length(nCount_RNA),y=sort(nCount_RNA))) +
           geom_point(size=0.5,alpha=0.5,color="#00BFC4") +
           labs(title=collection, x="Cell Rank", y="UMI per Cell") +
           ylim(0,100000)
  p4c <- ggplot(subset(data[[i]]@meta.data,Species%in%"Chimp"),
                aes(x=1:length(nFeature_RNA),y=sort(nFeature_RNA))) +
           geom_point(size=0.5,alpha=0.5,color="#F8766D") +
           labs(title=collection, x="Cell Rank", y="Genes per Cell") +
           ylim(0,10000)
  p4d <- ggplot(subset(data[[i]]@meta.data,Species%in%"Human"),
                aes(x=1:length(nFeature_RNA),y=sort(nFeature_RNA))) +
           geom_point(size=0.5,alpha=0.5,color="#00BFC4") +
           labs(title=collection, x="Cell Rank", y="Genes per Cell") +
           ylim(0,10000)
  
  #look at genes instead of cells
  #umis per gene & rarity of genes (cells expressing genes)
  #thresholds: umi per gene > 1
  df <- data.frame(
    counts_per_gene_chimp <- Matrix::rowSums(data.filter[[i]]@assays$RNA@counts[,which(data.filter[[i]]@meta.data$Species=="Chimp")]),
    counts_per_gene_human <- Matrix::rowSums(data.filter[[i]]@assays$RNA@counts[,which(data.filter[[i]]@meta.data$Species=="Human")]),
    cells_per_gene_chimp <- Matrix::rowSums(data.filter[[i]]@assays$RNA@counts[,which(data.filter[[i]]@meta.data$Species=="Chimp")]>0),
    cells_per_gene_human <- Matrix::rowSums(data.filter[[i]]@assays$RNA@counts[,which(data.filter[[i]]@meta.data$Species=="Human")]>0)
  )
  rm(counts_per_gene_chimp,counts_per_gene_human,cells_per_gene_chimp,cells_per_gene_human)
  colnames(df) <- c("counts_per_gene_chimp","counts_per_gene_human","cells_per_gene_chimp","cells_per_gene_human")

  print(paste0("Total Genes: ",length(rownames(data.filter[[i]]@assays$RNA@counts))))
  print("Genes with at least one UMI (Chimp, Human, and Total)")
  print(table(Matrix::rowSums(data.filter[[i]]@assays$RNA@counts[,which(data.filter[[i]]@meta.data$Species=="Chimp")])>0))
  print(table(Matrix::rowSums(data.filter[[i]]@assays$RNA@counts[,which(data.filter[[i]]@meta.data$Species=="Human")])>0))
  print(table(Matrix::rowSums(data.filter[[i]]@assays$RNA@counts)>0))
  
  p5a <- ggplot(df, aes(x=log10(counts_per_gene_chimp+1))) +
           geom_histogram(color="#F8766D", fill="white", alpha=0.1, binwidth=0.05) +
           ylim(0,6000) +
           geom_vline(xintercept=log10(1.5), linetype="dashed") +
           labs(title=collection, x="Log10 UMIs per Gene (Chimp)", y="Count")
  p5b <- ggplot(df, aes(x=log10(counts_per_gene_human+1))) +
           geom_histogram(color="#00BFC4", fill="white", alpha=0.1, binwidth=0.05) +
           ylim(0,6000) +
           geom_vline(xintercept=log10(1.5), linetype="dashed") +
           labs(title=collection, x="Log10 UMIs per Gene (Human)", y="Count")
  p5c <- ggplot(df, aes(x=log10(cells_per_gene_chimp+1))) +
           geom_histogram(color="#F8766D", fill="white", alpha=0.1, binwidth=0.05) +
           ylim(0,6000) +
           geom_vline(xintercept=log10(1.5), linetype="dashed") +
           labs(title=collection,x="Log10 Cells per Gene (Chimp)",y="Count")
  p5d <- ggplot(df, aes(x=log10(cells_per_gene_human+1))) +
           geom_histogram(color="#00BFC4", fill="white", alpha=0.1, binwidth=0.05) +
           ylim(0,6000) +
           geom_vline(xintercept=log10(1.5), linetype="dashed") +
           labs(title=collection,x="Log10 Cells per Gene (Human)",y="Count")

  grid.arrange(grobs=list(p1a,p1b,p1c,p1d,
                          p2a,p2b,p2c,p2d,
                          p3a,p3b,p3c,p3d,
                          p4a,p4b,p4c,p4d,
                          p5a,p5b,p5c,p5d), ncol=4)
  
  rm(p1a,p1b,p1c,p1d, p2a,p2b,p2c,p2d, p3a,p3b,p3c,p3d, p4a,p4b,p4c,p4d, p5a,p5b,p5c,p5d)
  
}

dev.off()

```
