---
title: "10x-data-integrate"
author: "Genevieve Housman"
date: "August 2020"
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## 10X Data Cellranger Analysis - Integrate Data

For each sample, normalization, variance stabilization, and regression of unwanted variation were performed. Following this, samples were integrated using shared highly variable genes.

For the first...

Another question was how should the data be subsetted before normalization. The following subsets were assessed:

* separate species and cell types (n=6) (EX: Human iPSCs)
* separate human-chimp pairs and cell types (n=21) (EX: H1+C1 iPSCs)
* separate individuals and cell types (n=42) (EX: H1 iPSCs)
* separate species but keep all cell types together (n=2) (EX: Human iPSCs+MSCs+Osteoblasts)
* separate human-chimp pairs but keep all cell types together (n=7) (EX: H1+C1 iPSCs+MSCs+Osteoblasts)
* separate individuals but keep all cell types together (n=14) (EX: H1 iPSCs+MSCs+Osteoblasts)

What I used:

* separate individuals and cell types (n=42) (EX: H1 iPSCs)
* separate individuals but keep all cell types together

**Step 1 - Normalization**

In Seurat the main normalization methods are:

1. Log Normalization
    + Within a cell, gene counts are divided by the total counts, multiplied by a scale.factor (10000), and natural-log transformed using log1p.
2. SCTransform
    + Alternative workflow that perfoms log normalization, variable feature identification, and data scaling in one command.

Because data are more difficult to manipulate with SCTransform, standard log normalization is used here.
Also, because 

**Step 2 - Calculate Cell Cycle Score**

**Step 3 - Variance Stabilization**

**Step 4 - Regression of Unwanted Variation**

**Step 5 - Data Integration**





**References**

* https://hbctraining.github.io/scRNA-seq/lessons/06_SC_SCT_and_integration.html
* See the following scripts for previous trial runs:
  + ./analysis/v1/v1-10x-data-normalize-log.Rmd
  + ./analysis/v1/v1-10x-data-normalize-sct.Rmd


```{bash, eval=FALSE, echo=TRUE}

### PROBABLY DO NOT NEED TO RUN ON CLUSTER THIS WAY

sinteractive --mem=24g --time=12:00:00 --partition=broadwl
module load R/3.6.1
export PYTHONUSERBASE=$HOME/.local/python
R
```


### Normalize Data & Find Variable Features

```{r prepare scRNA data for merging/integration - separate individuals and cell types}

#BASHSCRIPT: prep.l_indv-cell.sh (or prep.filter.l_indv-cell.sh / prep.filterMT.l_indv-cell.sh)

##!/bin/bash
##SBATCH --job-name=prep.l_indv-cell
##SBATCH --output=prep.l_indv-cell.out
##SBATCH --error=prep.l_indv-cell.err
##SBATCH --time=03:00:00
##SBATCH --partition=bigmem2
##SBATCH --nodes=1
##SBATCH --ntasks-per-node=8
##SBATCH --mem-per-cpu=25600
#module load R/3.5.1
#export PYTHONUSERBASE=$HOME/.local/python
#R CMD BATCH --no-save --no-restore prep.l_indv-cell.R prep.l_indv-cell.out

#RSCRIPT: prep.l_indv-cell.R (or prep.filter.l_indv-cell.R / prep.filterMT.l_indv-cell.R)

#Load libraries
library(Seurat)
library(dplyr)
library(stringi)
library(stringr)
library(ggplot2)
library(colorspace)
library(RColorBrewer)

#Define main directory
dir <- '/project2/gilad/ghousman/skeletal-human-chimp/'

#Load batch info
batch <- read.csv(file=paste0(dir,'human-chimp-skeletal-scRNA/data/scrna-batch.csv'), header=TRUE, sep=",")

#Read in files
data <- readRDS(paste0(dir,"/human-chimp-skeletal-scRNA/data/cellranger-data-full/data.rds"))

#Separate human and chimp data within each collection into separate objects
objects <- list()
idx <- 1
for (i in 1:length(batch$Sample_Name_at_Core))
{
  obj <- subset(data[[i]], species=='Human')
  objects[[idx]] <- CreateSeuratObject(obj@assays$RNA@counts, meta.data=obj@meta.data)
  idx <- idx+1
  obj <- subset(data[[i]], species=='Chimp')
  objects[[idx]] <- CreateSeuratObject(obj@assays$RNA@counts, meta.data=obj@meta.data)
  idx <- idx+1
}
rm(data,obj)

#Perform log normalization on each individual and cell type
for (i in 1:length(objects)) {
  print(i)
  objects[[i]] <- NormalizeData(objects[[i]], verbose=FALSE)
  objects[[i]] <- FindVariableFeatures(objects[[i]], selection.method="vst", nfeatures=3000, verbose=FALSE)
}

#Save data
saveRDS(objects, file=paste0(dir,"/human-chimp-skeletal-scRNA/data/cellranger-data-full/data.indv-cell.log.rds"))

```

```{r prepare scRNA data for merging/integration - separate individuals but keep all cell types together}

#BASHSCRIPT: prep.l_indv.sh  (or prep.filter.l_indv.sh / prep.filterMT.l_indv.sh)

##!/bin/bash
##SBATCH --job-name=prep.l_indv
##SBATCH --output=prep.l_indv.out
##SBATCH --error=prep.l_indv.err
##SBATCH --time=03:00:00
##SBATCH --partition=broadwl
##SBATCH --nodes=1
##SBATCH --ntasks-per-node=8
##SBATCH --mem-per-cpu=6400
#module load R/3.5.1
#export PYTHONUSERBASE=$HOME/.local/python
#R CMD BATCH --no-save --no-restore prep.l_indv.R prep.l_indv.out

#RSCRIPT: prep.l_indv.R (or prep.filter.l_indv.R / prep.filterMT.l_indv.R)

#Load libraries
library(Seurat)
library(dplyr)
library(stringi)
library(stringr)
library(ggplot2)
library(colorspace)
library(RColorBrewer)

#Define main directory
dir <- '/project2/gilad/ghousman/skeletal-human-chimp/'

#Load batch info
batch <- read.csv(file=paste0(dir,'human-chimp-skeletal-scRNA/data/scrna-batch.csv'), header=TRUE, sep=",")

#Read in files
data <- readRDS(paste0(dir,"/human-chimp-skeletal-scRNA/data/cellranger-data-full/data.rds"))

#Separate human and chimp data within each collection into separate objects
objects <- list()
idx <- 1
for (i in 1:length(batch$Sample_Name_at_Core))
{
  obj <- subset(data[[i]], species=='Human')
  objects[[idx]] <- CreateSeuratObject(obj@assays$RNA@counts, meta.data=obj@meta.data)
  idx <- idx+1
  obj <- subset(data[[i]], species=='Chimp')
  objects[[idx]] <- CreateSeuratObject(obj@assays$RNA@counts, meta.data=obj@meta.data)
  idx <- idx+1
}
rm(data,obj)

#Merge all cell types in each individual
h1 <- merge(objects[[1]], y=c(objects[[3]],objects[[5]]),add.cell.ids=c("H1.I","H1.M","H1.O"))
h1r2 <- merge(objects[[7]], y=c(objects[[9]],objects[[11]]),add.cell.ids=c("H1r2.I","H1r2.M","H1r2.O"))
h2 <- merge(objects[[13]], y=c(objects[[15]],objects[[17]]),add.cell.ids=c("H2.I","H2.M","H2.O"))
h3 <- merge(objects[[19]], y=c(objects[[21]],objects[[23]]),add.cell.ids=c("H3.I","H3.M","H3.O"))
h4 <- merge(objects[[25]], y=c(objects[[27]],objects[[29]]),add.cell.ids=c("H4.I","H4.M","H4.O"))
h5 <- merge(objects[[31]], y=c(objects[[33]],objects[[35]]),add.cell.ids=c("H5.I","H5.M","H5.O"))
h6 <- merge(objects[[37]], y=c(objects[[39]],objects[[41]]),add.cell.ids=c("H6.I","H6.M","H6.O"))
c1 <- merge(objects[[2]], y=c(objects[[4]],objects[[6]]),add.cell.ids=c("C1.I","C1.M","C1.O"))
c1r2 <- merge(objects[[8]], y=c(objects[[10]],objects[[12]]),add.cell.ids=c("C1r2.I","C1r2.M","C1r2.O"))
c2 <- merge(objects[[14]], y=c(objects[[16]],objects[[18]]),add.cell.ids=c("C2.I","C2.M","C2.O"))
c3 <- merge(objects[[20]], y=c(objects[[22]],objects[[24]]),add.cell.ids=c("C3.I","C3.M","C3.O"))
c4 <- merge(objects[[26]], y=c(objects[[28]],objects[[30]]),add.cell.ids=c("C4.I","C4.M","C4.O"))
c5 <- merge(objects[[32]], y=c(objects[[34]],objects[[36]]),add.cell.ids=c("C5.I","C5.M","C5.O"))
c6 <- merge(objects[[38]], y=c(objects[[40]],objects[[42]]),add.cell.ids=c("C6.I","C6.M","C6.O"))

objects <- list(h1,h1r2,h2,h3,h4,h5,h6,c1,c1r2,c2,c3,c4,c5,c6)

#Perform log normalization on each collection
for (i in 1:length(objects)) {
  print(i)
  objects[[i]] <- NormalizeData(objects[[i]], verbose=FALSE)
  objects[[i]] <- FindVariableFeatures(objects[[i]], selection.method="vst", nfeatures=3000, verbose=FALSE)
}

#Save data
saveRDS(objects, file=paste0(dir,"/human-chimp-skeletal-scRNA/data/cellranger-data-full/data.indv.log.rds"))

```

### Cell Cycle Scoring

References

* https://hbctraining.github.io/scRNA-seq/lessons/cell_cycle_scoring.html

```{r cell cycle}

#cc_file <- getURL("https://raw.githubusercontent.com/hbc/tinyatlas/master/cell_cycle/Homo_sapiens.csv") 
#cell_cycle_genes <- read.csv(text = cc_file)

#cell cycle markers (Tirosh et al, 2015) - loaded with seurat
s.genes <- cc.genes$s.genes
g2m.genes <- cc.genes$g2m.genes

#score cell cycle phases
indv <- CellCycleScoring(indv, s.features=s.genes, g2m.features=g2m.genes, set.ident=TRUE)

#view cell cycle scores and phase assignments
head(indv[[]])

#plot cell cycle scores
CombinePlots(plots=list((DimPlot(indv,
                                 group.by="Phase",
                                 reduction="pca",
                                 dims=c(1,2)) +
                           labs(x="PC1",y="PC2")),
                        (DimPlot(indv,
                                 group.by="Phase",
                                 reduction="pca",
                                 dims=c(3,4)) +
                           labs(x="PC3",y="PC4")),
                        (DimPlot(indv,
                                 group.by="Phase",
                                 reduction="pca",
                                 dims=c(5,6)) +
                           labs(x="PC5",y="PC6")),
                        (DimPlot(indv,
                                 group.by="Phase",
                                 reduction="pca",
                                 dims=c(7,8)) +
                           labs(x="PC7",y="PC8")),
                        (DimPlot(indv,
                                 group.by="Phase",
                                 reduction="pca",
                                 dims=c(9,10)) +
                           labs(x="PC9",y="PC10"))), ncol=5)
CombinePlots(plots=list((DimPlot(subset(indv,subset=species=="Human"),
                                 group.by="Phase",
                                 reduction="umap") +
                           labs(x="UMAP1",y="UMAP2")),
                        (DimPlot(subset(indv,subset=species=="Chimp"),
                                 group.by="Phase",
                                 reduction="umap") +
                           labs(x="UMAP1",y="UMAP2"))), ncol=1)
CombinePlots(plots=list((DimPlot(subset(indv,subset=Phase=="G1"),
                                 group.by="Cell.Type",
                                 reduction="umap",
                                 cols=c("pink","mediumorchid","dodgerblue")) +
                           labs(x="UMAP1",y="UMAP2")),
                        (DimPlot(subset(indv,subset=Phase=="G2M"),
                                 group.by="Cell.Type",
                                 reduction="umap",
                                 cols=c("pink","mediumorchid","dodgerblue")) +
                           labs(x="UMAP1",y="UMAP2")),
                        (DimPlot(subset(indv,subset=Phase=="S"),
                                 group.by="Cell.Type",
                                 reduction="umap",
                                 cols=c("pink","mediumorchid","dodgerblue")) +
                           labs(x="UMAP1",y="UMAP2"))), ncol=3)


```

### Data Integration

What I used:

* separate individuals and cells types
* separate individuals but keep all cell types together
* merge data without integration

...something about merging datasets using the intersect of the 6000 most variable genes in humans and chimpanzees

```{r integrate scRNA data - separate individuals and cells types}

#BASHSCRIPT: rpca.l_indv-cell.sh (or rpca.filter.l_indv-cell.sh / rpca.l.filterMT.l_indv-cell.sh)

##!/bin/bash
##SBATCH --job-name=rpca.l_indv-cell
##SBATCH --output=rpca.l_indv-cell.out
##SBATCH --error=rpca.l_indv-cell.err
##SBATCH --time=03:00:00
##SBATCH --partition=broadwl
##SBATCH --nodes=1
##SBATCH --ntasks-per-node=8
##SBATCH --mem-per-cpu=6400
#module load R/3.5.1
#export PYTHONUSERBASE=$HOME/.local/python
#R CMD BATCH --no-save --no-restore rpca.l_indv-cell.R rpca.l_indv-cell.out

#RSCRIPT: rpca.l_indv-cell.R (or rpca.filter.l_indv-cell.R / rpca.filterMT.l_indv-cell.R)

#Load libraries
library(Seurat)
library(dplyr)
library(stringi)
library(stringr)
library(ggplot2)
library(colorspace)
library(RColorBrewer)

#Define main directory
dir <- '/project2/gilad/ghousman/skeletal-human-chimp/'

#Load batch info
batch <- read.csv(file=paste0(dir,'human-chimp-skeletal-scRNA/data/scrna-batch.csv'), header=TRUE, sep=",")

#Load objects
objects <- readRDS(paste0(dir,"/human-chimp-skeletal-scRNA/data/cellranger-data-full/v1/data.indv-cell.log.rds"))

#Define list subset
#Select features for downstream integration (keeping at 1000 for now)
#Identify anchors (used references + RPCA reduction method) #ALL OTHER METHODS CRASH
#Integrate datasets
#Reduce dimensionality
#Save data

#combine ipscs from all samples
obj.ipsc  <- list(objects[[1]],objects[[2]],objects[[7]],objects[[8]],objects[[13]],objects[[14]],objects[[19]],objects[[20]],
                  objects[[25]],objects[[26]],objects[[31]],objects[[32]],objects[[37]],objects[[38]])
obj.ipsc.features <- SelectIntegrationFeatures(object.list=obj.ipsc, nfeatures=1000)
reference_dataset <- c(3:4) #H1C1-r2 differentiated the best
obj.ipsc <- lapply(X=obj.ipsc, FUN=function(x) {
    x <- ScaleData(x, features=obj.ipsc.features, verbose=FALSE)
    x <- RunPCA(x, features=obj.ipsc.features, verbose=FALSE)
})
obj.ipsc.anchors <- FindIntegrationAnchors(object.list=obj.ipsc, normalization.method="LogNormalize", anchor.features=obj.ipsc.features,
                                           reference=reference_dataset, reduction="rpca")
int.ipsc <- IntegrateData(anchorset=obj.ipsc.anchors, normalization.method="LogNormalize")
int.ipsc <- ScaleData(int.ipsc, verbose=FALSE)
int.ipsc <- RunPCA(object=int.ipsc, npcs=100, verbose=FALSE)
pva <- int.ipsc@reductions$pca@stdev^2/int.ipsc@reductions$pca@misc$total.variance
ndim <- length(which(pva>=0.001)) #keep all dims that explaim more than 0.1% of variance
print(ndim)
int.ipsc <- RunUMAP(int.ipsc, dims=1:100)
saveRDS(int.ipsc, file=paste0(dir,"/human-chimp-skeletal-scRNA/data/cellranger-data-full/data.indv-cell.log.int.ipsc.rds"))
rm(obj.ipsc,int.ipsc)

#combine mscs from all samples
obj.msc   <- list(objects[[3]],objects[[4]],objects[[9]],objects[[10]],objects[[15]],objects[[16]],objects[[21]],objects[[22]],
                  objects[[27]],objects[[28]],objects[[33]],objects[[34]],objects[[39]],objects[[40]])
obj.msc.features <- SelectIntegrationFeatures(object.list=obj.msc, nfeatures=1000)
reference_dataset <- c(3:4) #H1C1-r2 differentiated the best
obj.msc <- lapply(X=obj.msc, FUN=function(x) {
    x <- ScaleData(x, features=obj.msc.features, verbose=FALSE)
    x <- RunPCA(x, features=obj.msc.features, verbose=FALSE)
})
obj.msc.anchors <- FindIntegrationAnchors(object.list=obj.msc, normalization.method="LogNormalize", anchor.features=obj.msc.features,
                                           reference=reference_dataset, reduction="rpca")
int.msc <- IntegrateData(anchorset=obj.msc.anchors, normalization.method="LogNormalize")
int.msc <- ScaleData(int.msc, verbose=FALSE)
int.msc <- RunPCA(object=int.msc, npcs=100, verbose=FALSE)
pva <- int.msc@reductions$pca@stdev^2/int.msc@reductions$pca@misc$total.variance
ndim <- length(which(pva>=0.001)) #keep all dims that explaim more than 0.1% of variance
print(ndim)
int.msc <- RunUMAP(int.msc, dims=1:100)
saveRDS(int.msc, file=paste0(dir,"/human-chimp-skeletal-scRNA/data/cellranger-data-full/data.indv-cell.log.int.msc.rds"))
rm(obj.msc,int.msc)

#combine osteoblasts from all samples
obj.osteo <- list(objects[[5]],objects[[6]],objects[[11]],objects[[12]],objects[[17]],objects[[18]],objects[[23]],objects[[24]],
                  objects[[29]],objects[[30]],objects[[35]],objects[[36]],objects[[41]],objects[[42]])
obj.osteo.features <- SelectIntegrationFeatures(object.list=obj.osteo, nfeatures=1000)
reference_dataset <- c(3:4) #H1C1-r2 differentiated the best
obj.osteo <- lapply(X=obj.osteo, FUN=function(x) {
    x <- ScaleData(x, features=obj.osteo.features, verbose=FALSE)
    x <- RunPCA(x, features=obj.osteo.features, verbose=FALSE)
})
obj.osteo.anchors <- FindIntegrationAnchors(object.list=obj.osteo, normalization.method="LogNormalize", anchor.features=obj.osteo.features,
                                           reference=reference_dataset, reduction="rpca")
int.osteo <- IntegrateData(anchorset=obj.osteo.anchors, normalization.method="LogNormalize")
int.osteo <- ScaleData(int.osteo, verbose=FALSE)
int.osteo <- RunPCA(object=int.osteo, npcs=100, verbose=FALSE)
pva <- int.osteo@reductions$pca@stdev^2/int.osteo@reductions$pca@misc$total.variance
ndim <- length(which(pva>=0.001)) #keep all dims that explaim more than 0.1% of variance
print(ndim)
int.osteo <- RunUMAP(int.osteo, dims=1:100)
saveRDS(int.osteo, file=paste0(dir,"/human-chimp-skeletal-scRNA/data/cellranger-data-full/data.indv-cell.log.int.osteo.rds"))
rm(obj.osteo,int.osteo)

```

```{r integrate scRNA data - separate individuals but keep all cell types together}

#BASHSCRIPT: rpca.l_indv.sh (or rpca.filter.l_indv.sh / rpca.filterMT.l_indv.sh)

##!/bin/bash
##SBATCH --job-name=rpca.l_indv
##SBATCH --output=rpca.l_indv.out
##SBATCH --error=rpca.l_indv.err
##SBATCH --time=03:00:00
##SBATCH --partition=bigmem2
##SBATCH --nodes=1
##SBATCH --ntasks-per-node=8
##SBATCH --mem-per-cpu=12800
#module load R/3.5.1
#export PYTHONUSERBASE=$HOME/.local/python
#R CMD BATCH --no-save --no-restore rpca.l_indv.R rpca.l_indv.out

#RSCRIPT: rpca.l_indv.R (or rpca.filter.l_indv.R / rpca.filterMT.l_indv.R)

#Load libraries
library(Seurat)
library(dplyr)
library(stringi)
library(stringr)
library(ggplot2)
library(colorspace)
library(RColorBrewer)

#Define main directory
dir <- '/project2/gilad/ghousman/skeletal-human-chimp/'

#Load batch info
batch <- read.csv(file=paste0(dir,'human-chimp-skeletal-scRNA/data/scrna-batch.csv'), header=TRUE, sep=",")

#Load objects
objects <- readRDS(paste0(dir,"/human-chimp-skeletal-scRNA/data/cellranger-data-full/data.indv.log.rds"))

#Select features for downstream integration (keeping at 1000 for now)
#Identify anchors (used references + RPCA reduction method) #ALL OTHER METHODS CRASH
objects.features <- SelectIntegrationFeatures(object.list=objects, nfeatures=1000)
reference_dataset <- c(2,9) #make H1-r2 and C1-r2 the references
objects <- lapply(X=objects, FUN=function(x) {
    x <- ScaleData(x, features=objects.features, verbose=FALSE)
    x <- RunPCA(x, features=objects.features, verbose=FALSE)
})
objects.anchors <- FindIntegrationAnchors(object.list=objects, normalization.method="LogNormalize", anchor.features=objects.features,
                                          reference=reference_dataset, reduction="rpca")

#Integrate datasets
integrate <- IntegrateData(anchorset=objects.anchors, normalization.method="LogNormalize")
integrate <- ScaleData(integrate, verbose=FALSE)

#Reduce dimensionality
integrate <- RunPCA(object=integrate, npcs=100, verbose=FALSE)
pva <- integrate@reductions$pca@stdev^2/integrate@reductions$pca@misc$total.variance
ndim <- length(which(pva>=0.001)) #keep all dims that explaim more than 0.1% of variance
print(ndim)
integrate <- RunUMAP(integrate, dims=1:100)

#Save data
saveRDS(integrate, file=paste0(dir,"/human-chimp-skeletal-scRNA/data/cellranger-data-full/data.indv.log.int.rds"))
rm(objects,integrate)

```

```{r examine batch effects - merged data vs. integrated data}

#Load libraries
library(Seurat)
library(dplyr)
library(stringi)
library(stringr)
library(ggplot2)
library(colorspace)
library(RColorBrewer)

#Define main directory
dir <- '/project2/gilad/ghousman/skeletal-human-chimp/'

#Load batch info
batch <- read.csv(file=paste0(dir,'human-chimp-skeletal-scRNA/data/scrna-batch.csv'), header=TRUE, sep=",")

#Read and Visualize data (batch of interest: species, Pair, Individual, Cell.Type, Replicate, MSCmedia)

#combine ipscs from all samples
data.ipsc <- readRDS(paste0(dir,"/human-chimp-skeletal-scRNA/data/cellranger-data-full/data.merge.ipsc.rds"))             #merge
spps.ipsc <- readRDS(paste0(dir,"/human-chimp-skeletal-scRNA/data/cellranger-data-full/data.spps-cell.log.int.ipsc.rds")) #integrate across species
pair.ipsc <- readRDS(paste0(dir,"/human-chimp-skeletal-scRNA/data/cellranger-data-full/data.pair-cell.log.int.ipsc.rds")) #integrate across human-chimp pairs
indv.ipsc <- readRDS(paste0(dir,"/human-chimp-skeletal-scRNA/data/cellranger-data-full/data.indv-cell.log.int.ipsc.rds")) #integrate across individuals

CombinePlots(plots=list(DimPlot(data.ipsc,group.by="species"),
                        DimPlot(data.ipsc,group.by="Pair"),
                        DimPlot(data.ipsc,group.by="Individual"),
                        DimPlot(data.ipsc,group.by="Cell.Type"),
                        DimPlot(data.ipsc,group.by="Replicate"),
                        DimPlot(data.ipsc,group.by="MSCmedia"),

                        DimPlot(spps.ipsc,group.by="species"),
                        DimPlot(spps.ipsc,group.by="Pair"),
                        DimPlot(spps.ipsc,group.by="Individual"),
                        DimPlot(spps.ipsc,group.by="Cell.Type"),
                        DimPlot(spps.ipsc,group.by="Replicate"),
                        DimPlot(spps.ipsc,group.by="MSCmedia"),
                        
                        DimPlot(pair.ipsc,group.by="species"),
                        DimPlot(pair.ipsc,group.by="Pair"),
                        DimPlot(pair.ipsc,group.by="Individual"),
                        DimPlot(pair.ipsc,group.by="Cell.Type"),
                        DimPlot(pair.ipsc,group.by="Replicate"),
                        DimPlot(pair.ipsc,group.by="MSCmedia"),
                        
                        DimPlot(indv.ipsc,group.by="species"),
                        DimPlot(indv.ipsc,group.by="Pair"),
                        DimPlot(indv.ipsc,group.by="Individual"),
                        DimPlot(indv.ipsc,group.by="Cell.Type"),
                        DimPlot(indv.ipsc,group.by="Replicate"),
                        DimPlot(indv.ipsc,group.by="MSCmedia")), ncol=5)

CombinePlots(plots=list(DimPlot(data.ipsc,group.by="species",reduction='pca'),
                        DimPlot(data.ipsc,group.by="Collection",reduction='pca'),
                        DimPlot(data.ipsc,group.by="Cell.Type",reduction='pca'),
                        DimPlot(data.ipsc,group.by="Samples",reduction='pca'),
                        DimPlot(data.ipsc,group.by="Replicate",reduction='pca'),

                        DimPlot(spps.ipsc,group.by="species",reduction='pca'),
                        DimPlot(spps.ipsc,group.by="Pair",reduction='pca'),
                        DimPlot(spps.ipsc,group.by="Individual",reduction='pca'),
                        DimPlot(spps.ipsc,group.by="Cell.Type",reduction='pca'),
                        DimPlot(spps.ipsc,group.by="Replicate",reduction='pca'),
                        DimPlot(spps.ipsc,group.by="MSCmedia",reduction='pca'),
                        
                        DimPlot(pair.ipsc,group.by="species",reduction='pca'),
                        DimPlot(pair.ipsc,group.by="Pair",reduction='pca'),
                        DimPlot(pair.ipsc,group.by="Individual",reduction='pca'),
                        DimPlot(pair.ipsc,group.by="Cell.Type",reduction='pca'),
                        DimPlot(pair.ipsc,group.by="Replicate",reduction='pca'),
                        DimPlot(pair.ipsc,group.by="MSCmedia",reduction='pca'),
                        
                        DimPlot(indv.ipsc,group.by="species",reduction='pca'),
                        DimPlot(indv.ipsc,group.by="Pair",reduction='pca'),
                        DimPlot(indv.ipsc,group.by="Individual",reduction='pca'),
                        DimPlot(indv.ipsc,group.by="Cell.Type",reduction='pca'),
                        DimPlot(indv.ipsc,group.by="Replicate",reduction='pca'),
                        DimPlot(indv.ipsc,group.by="MSCmedia",reduction='pca')), ncol=6)

CombinePlots(plots=list(VlnPlot(data.ipsc,'nCount_RNA',group.by='species'),
                        VlnPlot(spps.ipsc,'nCount_RNA',group.by='species'),
                        VlnPlot(pair.ipsc,'nCount_RNA',group.by='species'),
                        VlnPlot(indv.ipsc,'nCount_RNA',group.by='species'),
                        
                        VlnPlot(data.ipsc,'nFeature_RNA',group.by='species'),
                        VlnPlot(spps.ipsc,'nFeature_RNA',group.by='species'),
                        VlnPlot(pair.ipsc,'nFeature_RNA',group.by='species'),
                        VlnPlot(indv.ipsc,'nFeature_RNA',group.by='species')), ncol=4)

rm(data.ipsc, spps.ipsc, pair.ipsc, indv.ipsc)

#data.ipsc:   species and batches separated           (PC/UMAP separates species, PC/UMAP separates batches)
#spps.ipsc:   species integrated / batches separated  (PC/UMAP separates species)
#pair.ipsc:   species separated / batches integrated  (PC/UMAP separates species)
#indv.ipsc:   species and batches integrated
#NOTES:       UMI and gene counts are higher in humans than chimps


#combine mscs from all samples
data.msc <- readRDS(paste0(dir,"/human-chimp-skeletal-scRNA/data/cellranger-data-full/data.merge.msc.rds"))             #merge
spps.msc <- readRDS(paste0(dir,"/human-chimp-skeletal-scRNA/data/cellranger-data-full/data.spps-cell.log.int.msc.rds")) #integrate across species
pair.msc <- readRDS(paste0(dir,"/human-chimp-skeletal-scRNA/data/cellranger-data-full/data.pair-cell.log.int.msc.rds")) #integrate across human-chimp pairs
indv.msc <- readRDS(paste0(dir,"/human-chimp-skeletal-scRNA/data/cellranger-data-full/data.indv-cell.log.int.msc.rds")) #integrate across individuals

CombinePlots(plots=list(DimPlot(data.msc,group.by="species"),
                        DimPlot(data.msc,group.by="Pair"),
                        DimPlot(data.msc,group.by="Individual"),
                        DimPlot(data.msc,group.by="Cell.Type"),
                        DimPlot(data.msc,group.by="Replicate"),
                        DimPlot(data.msc,group.by="MSCmedia"),

                        DimPlot(spps.msc,group.by="species"),
                        DimPlot(spps.msc,group.by="Pair"),
                        DimPlot(spps.msc,group.by="Individual"),
                        DimPlot(spps.msc,group.by="Cell.Type"),
                        DimPlot(spps.msc,group.by="Replicate"),
                        DimPlot(spps.msc,group.by="MSCmedia"),
                        
                        DimPlot(pair.msc,group.by="species"),
                        DimPlot(pair.msc,group.by="Pair"),
                        DimPlot(pair.msc,group.by="Individual"),
                        DimPlot(pair.msc,group.by="Cell.Type"),
                        DimPlot(pair.msc,group.by="Replicate"),
                        DimPlot(pair.msc,group.by="MSCmedia"),
                        
                        DimPlot(indv.msc,group.by="species"),
                        DimPlot(indv.msc,group.by="Pair"),
                        DimPlot(indv.msc,group.by="Individual"),
                        DimPlot(indv.msc,group.by="Cell.Type"),
                        DimPlot(indv.msc,group.by="Replicate"),
                        DimPlot(indv.msc,group.by="MSCmedia")), ncol=5)

CombinePlots(plots=list(DimPlot(data.msc,group.by="species",reduction='pca'),
                        DimPlot(data.msc,group.by="Collection",reduction='pca'),
                        DimPlot(data.msc,group.by="Cell.Type",reduction='pca'),
                        DimPlot(data.msc,group.by="Samples",reduction='pca'),
                        DimPlot(data.msc,group.by="Replicate",reduction='pca'),

                        DimPlot(spps.msc,group.by="species",reduction='pca'),
                        DimPlot(spps.msc,group.by="Pair",reduction='pca'),
                        DimPlot(spps.msc,group.by="Individual",reduction='pca'),
                        DimPlot(spps.msc,group.by="Cell.Type",reduction='pca'),
                        DimPlot(spps.msc,group.by="Replicate",reduction='pca'),
                        DimPlot(spps.msc,group.by="MSCmedia",reduction='pca'),
                        
                        DimPlot(pair.msc,group.by="species",reduction='pca'),
                        DimPlot(pair.msc,group.by="Pair",reduction='pca'),
                        DimPlot(pair.msc,group.by="Individual",reduction='pca'),
                        DimPlot(pair.msc,group.by="Cell.Type",reduction='pca'),
                        DimPlot(pair.msc,group.by="Replicate",reduction='pca'),
                        DimPlot(pair.msc,group.by="MSCmedia",reduction='pca'),
                        
                        DimPlot(indv.msc,group.by="species",reduction='pca'),
                        DimPlot(indv.msc,group.by="Pair",reduction='pca'),
                        DimPlot(indv.msc,group.by="Individual",reduction='pca'),
                        DimPlot(indv.msc,group.by="Cell.Type",reduction='pca'),
                        DimPlot(indv.msc,group.by="Replicate",reduction='pca'),
                        DimPlot(indv.msc,group.by="MSCmedia",reduction='pca')), ncol=6)

CombinePlots(plots=list(VlnPlot(data.msc,'nCount_RNA',group.by='species'),
                        VlnPlot(spps.msc,'nCount_RNA',group.by='species'),
                        VlnPlot(pair.msc,'nCount_RNA',group.by='species'),
                        VlnPlot(indv.msc,'nCount_RNA',group.by='species'),
                        
                        VlnPlot(data.msc,'nFeature_RNA',group.by='species'),
                        VlnPlot(spps.msc,'nFeature_RNA',group.by='species'),
                        VlnPlot(pair.msc,'nFeature_RNA',group.by='species'),
                        VlnPlot(indv.msc,'nFeature_RNA',group.by='species')), ncol=4)

rm(data.msc, spps.msc, pair.msc, indv.msc)

#data.msc:   species and batches separated           (PC/UMAP separates species, PC/UMAP separates batches)
#spps.msc:   species integrated / batches separated  (PC/UMAP separates species)
#pair.msc:   species separated / batches integrated  (PC/UMAP separates species)
#indv.msc:   species and batches integrated
#NOTES:      UMI and gene counts are higher in humans than chimps


#combine osteoblasts from all samples
data.osteo <- readRDS(paste0(dir,"/human-chimp-skeletal-scRNA/data/cellranger-data-full/data.merge.osteo.rds"))             #merge
spps.osteo <- readRDS(paste0(dir,"/human-chimp-skeletal-scRNA/data/cellranger-data-full/data.spps-cell.log.int.osteo.rds")) #integrate across species
pair.osteo <- readRDS(paste0(dir,"/human-chimp-skeletal-scRNA/data/cellranger-data-full/data.pair-cell.log.int.osteo.rds")) #integrate across human-chimp pairs
indv.osteo <- readRDS(paste0(dir,"/human-chimp-skeletal-scRNA/data/cellranger-data-full/data.indv-cell.log.int.osteo.rds")) #integrate across individuals

CombinePlots(plots=list(DimPlot(data.osteo,group.by="species"),
                        DimPlot(data.osteo,group.by="Pair"),
                        DimPlot(data.osteo,group.by="Individual"),
                        DimPlot(data.osteo,group.by="Cell.Type"),
                        DimPlot(data.osteo,group.by="Replicate"),
                        DimPlot(data.osteo,group.by="MSCmedia"),

                        DimPlot(spps.osteo,group.by="species"),
                        DimPlot(spps.osteo,group.by="Pair"),
                        DimPlot(spps.osteo,group.by="Individual"),
                        DimPlot(spps.osteo,group.by="Cell.Type"),
                        DimPlot(spps.osteo,group.by="Replicate"),
                        DimPlot(spps.osteo,group.by="MSCmedia"),
                        
                        DimPlot(pair.osteo,group.by="species"),
                        DimPlot(pair.osteo,group.by="Pair"),
                        DimPlot(pair.osteo,group.by="Individual"),
                        DimPlot(pair.osteo,group.by="Cell.Type"),
                        DimPlot(pair.osteo,group.by="Replicate"),
                        DimPlot(pair.osteo,group.by="MSCmedia"),
                        
                        DimPlot(indv.osteo,group.by="species"),
                        DimPlot(indv.osteo,group.by="Pair"),
                        DimPlot(indv.osteo,group.by="Individual"),
                        DimPlot(indv.osteo,group.by="Cell.Type"),
                        DimPlot(indv.osteo,group.by="Replicate"),
                        DimPlot(indv.osteo,group.by="MSCmedia")), ncol=5)

CombinePlots(plots=list(DimPlot(data.osteo,group.by="species",reduction='pca'),
                        DimPlot(data.osteo,group.by="Collection",reduction='pca'),
                        DimPlot(data.osteo,group.by="Cell.Type",reduction='pca'),
                        DimPlot(data.osteo,group.by="Samples",reduction='pca'),
                        DimPlot(data.osteo,group.by="Replicate",reduction='pca'),

                        DimPlot(spps.osteo,group.by="species",reduction='pca'),
                        DimPlot(spps.osteo,group.by="Pair",reduction='pca'),
                        DimPlot(spps.osteo,group.by="Individual",reduction='pca'),
                        DimPlot(spps.osteo,group.by="Cell.Type",reduction='pca'),
                        DimPlot(spps.osteo,group.by="Replicate",reduction='pca'),
                        DimPlot(spps.osteo,group.by="MSCmedia",reduction='pca'),
                        
                        DimPlot(pair.osteo,group.by="species",reduction='pca'),
                        DimPlot(pair.osteo,group.by="Pair",reduction='pca'),
                        DimPlot(pair.osteo,group.by="Individual",reduction='pca'),
                        DimPlot(pair.osteo,group.by="Cell.Type",reduction='pca'),
                        DimPlot(pair.osteo,group.by="Replicate",reduction='pca'),
                        DimPlot(pair.osteo,group.by="MSCmedia",reduction='pca'),
                        
                        DimPlot(indv.osteo,group.by="species",reduction='pca'),
                        DimPlot(indv.osteo,group.by="Pair",reduction='pca'),
                        DimPlot(indv.osteo,group.by="Individual",reduction='pca'),
                        DimPlot(indv.osteo,group.by="Cell.Type",reduction='pca'),
                        DimPlot(indv.osteo,group.by="Replicate",reduction='pca'),
                        DimPlot(indv.osteo,group.by="MSCmedia",reduction='pca')), ncol=6)

CombinePlots(plots=list(VlnPlot(data.osteo,'nCount_RNA',group.by='species'),
                        VlnPlot(spps.osteo,'nCount_RNA',group.by='species'),
                        VlnPlot(pair.osteo,'nCount_RNA',group.by='species'),
                        VlnPlot(indv.osteo,'nCount_RNA',group.by='species'),
                        
                        VlnPlot(data.osteo,'nFeature_RNA',group.by='species'),
                        VlnPlot(spps.osteo,'nFeature_RNA',group.by='species'),
                        VlnPlot(pair.osteo,'nFeature_RNA',group.by='species'),
                        VlnPlot(indv.osteo,'nFeature_RNA',group.by='species')), ncol=4)

rm(data.osteo, spps.osteo, pair.osteo, indv.osteo)

#data.osteo:   species and batches separated           (PC/UMAP separates species, PC/UMAP separates batches)
#spps.osteo:   species integrated / batches separated  (PC/UMAP separates species)
#pair.osteo:   species separated / batches integrated  (PC/UMAP separates species)
#indv.osteo:   species and batches integrated
#NOTES:        UMI and gene counts are higher in humans than chimps


#combine all cells from all samples
data <- readRDS(paste0(dir,"/human-chimp-skeletal-scRNA/data/cellranger-data-full/data.merge.rds"))        #merge
spps <- readRDS(paste0(dir,"/human-chimp-skeletal-scRNA/data/cellranger-data-full/data.spps.log.int.rds")) #integrate across species
pair <- readRDS(paste0(dir,"/human-chimp-skeletal-scRNA/data/cellranger-data-full/data.pair.log.int.rds")) #integrate across human-chimp pairs
indv <- readRDS(paste0(dir,"/human-chimp-skeletal-scRNA/data/cellranger-data-full/data.indv.log.int.rds")) #integrate across individuals

CombinePlots(plots=list(DimPlot(data,group.by="species"),
                        DimPlot(data,group.by="Pair"),
                        DimPlot(data,group.by="Individual"),
                        DimPlot(data,group.by="Cell.Type"),
                        DimPlot(data,group.by="Replicate"),
                        DimPlot(data,group.by="MSCmedia"),

                        DimPlot(spps,group.by="species"),
                        DimPlot(spps,group.by="Pair"),
                        DimPlot(spps,group.by="Individual"),
                        DimPlot(spps,group.by="Cell.Type"),
                        DimPlot(spps,group.by="Replicate"),
                        DimPlot(spps,group.by="MSCmedia"),
                        
                        DimPlot(pair,group.by="species"),
                        DimPlot(pair,group.by="Pair"),
                        DimPlot(pair,group.by="Individual"),
                        DimPlot(pair,group.by="Cell.Type"),
                        DimPlot(pair,group.by="Replicate"),
                        DimPlot(pair,group.by="MSCmedia"),
                        
                        DimPlot(indv,group.by="species"),
                        DimPlot(indv,group.by="Pair"),
                        DimPlot(indv,group.by="Individual"),
                        DimPlot(indv,group.by="Cell.Type"),
                        DimPlot(indv,group.by="Replicate"),
                        DimPlot(indv,group.by="MSCmedia")), ncol=5)

CombinePlots(plots=list(DimPlot(data,group.by="species",reduction='pca'),
                        DimPlot(data,group.by="Collection",reduction='pca'),
                        DimPlot(data,group.by="Cell.Type",reduction='pca'),
                        DimPlot(data,group.by="Samples",reduction='pca'),
                        DimPlot(data,group.by="Replicate",reduction='pca'),

                        DimPlot(spps,group.by="species",reduction='pca'),
                        DimPlot(spps,group.by="Pair",reduction='pca'),
                        DimPlot(spps,group.by="Individual",reduction='pca'),
                        DimPlot(spps,group.by="Cell.Type",reduction='pca'),
                        DimPlot(spps,group.by="Replicate",reduction='pca'),
                        DimPlot(spps,group.by="MSCmedia",reduction='pca'),
                        
                        DimPlot(pair,group.by="species",reduction='pca'),
                        DimPlot(pair,group.by="Pair",reduction='pca'),
                        DimPlot(pair,group.by="Individual",reduction='pca'),
                        DimPlot(pair,group.by="Cell.Type",reduction='pca'),
                        DimPlot(pair,group.by="Replicate",reduction='pca'),
                        DimPlot(pair,group.by="MSCmedia",reduction='pca'),
                        
                        DimPlot(indv,group.by="species",reduction='pca'),
                        DimPlot(indv,group.by="Pair",reduction='pca'),
                        DimPlot(indv,group.by="Individual",reduction='pca'),
                        DimPlot(indv,group.by="Cell.Type",reduction='pca'),
                        DimPlot(indv,group.by="Replicate",reduction='pca'),
                        DimPlot(indv,group.by="MSCmedia",reduction='pca')), ncol=6)

CombinePlots(plots=list(VlnPlot(data,'nCount_RNA',group.by='species'),
                        VlnPlot(spps,'nCount_RNA',group.by='species'),
                        VlnPlot(pair,'nCount_RNA',group.by='species'),
                        VlnPlot(indv,'nCount_RNA',group.by='species'),
                        
                        VlnPlot(data,'nFeature_RNA',group.by='species'),
                        VlnPlot(spps,'nFeature_RNA',group.by='species'),
                        VlnPlot(pair,'nFeature_RNA',group.by='species'),
                        VlnPlot(indv,'nFeature_RNA',group.by='species'),
                        VlnPlot(int,'nFeature_RNA',group.by='species')), ncol=4)

rm(data, spps, pair, indv)

#data:   species + cell type + batches separated             (PC/UMAP separates species, PC/UMAP separates cell types, PC/UMAP separates batches)
#spps:   cell type + batches separated / species integrated  (PC/UMAP separates cell types, PC/UMAP separates batches)
#pair:   species + cell type separated / batches integrated  (PC/UMAP separates species, PC/UMAP separates cell types)
#indv:   cell type separated / species + batches integrated  (PC/UMAP separates cell types)
#NOTES:  UMI and gene counts are higher in humans than chimps

```

